### 2020-04-29 main.tex
further work

\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage[left=2.00cm, right=2.00cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{amsmath}
\usepackage[colorlinks]{hyperref}
\usepackage[backend=biber]{biblatex}
\usepackage{eurosym}
\usepackage[dvipsnames]{xcolor} 
\usepackage{subcaption}
\usepackage{accents}
\usepackage[capitalise]{cleveref}

\addbibresource{main.bib}
\graphicspath{{figures/}}

\crefname{relation}{Rel.}{Rels.}
\creflabelformat{relation}{(#2#1#3)}
\crefname{constraint}{Constr.}{Constrs.}
\creflabelformat{constraint}{(#2#1#3)}

\setlength\parindent{8pt}

\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\newcommand{\note}[1]{\textcolor{Orange}{#1}}

\newcommand{\generation}[1][n]{g_{#1,s,t}}
\newcommand{\generationpotential}{\bar{g}_{n,s,t}}
\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
\newcommand{\capacityGeneration}{G_{n,s}}
\newcommand{\capacityFlow}{F_{\ell}}
\newcommand{\capexGeneration}{c_{n,s}}
\newcommand{\capexFlow}{c_{\ell}}
\newcommand{\opexGeneration}[1][n]{o_{#1,s}}
\newcommand{\demand}[1][n]{d_{#1,a,t}}
\newcommand{\demandnodal}[1][n]{d_{#1,t}}
\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
\newcommand{\utility}{U_{n,a,t}}
\newcommand{\incidence}[1][n]{K_{#1,\ell}}
\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
\newcommand{\ptdfEqual}[1][n]{\ptdf[#1]^\circ}
\newcommand{\slackflow}{k_{\ell}}
\newcommand{\slack}[1][n]{k_{#1}}
\newcommand{\Slack}{k_{m,n}}
\newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
\newcommand{\muuppergeneration}[1][n]{\bar{\mu}_{#1,s,t}}
\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
\newcommand{\flow}{f_{\ell,t}}
\newcommand{\cycle}{C_{\ell,c}}
\newcommand{\impedance}{x_\ell}
\newcommand{\cycleprice}{\lambda_{c,t}}
\newcommand{\injection}{p_{n,t}}
\newcommand{\netconsumption}{p^{-}_{n,t}}
\newcommand{\netproduction}{p^{+}_{n,t}}
\newcommand{\selfconsumption}{p^{\circ}_{n,t}}
\newcommand{\lagrangian}{\mathcal{L}}

\newcommand{\allocatePeer}[1][m \rightarrow n]{A_{#1,t}}
\newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
\newcommand{\allocateTransaction}[1][m \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocatePeerNormed}[1][m \rightarrow n]{\tilde{A}_{#1,t}}
\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}

\newcommand{\emission}[1][n]{e_{#1,s}}
\newcommand{\emissionPrice}{\mu_{\text{CO2}}}
\newcommand{\megawatthour}{MWh$_\text{el}$}
\newcommand{\totalcost}{\mathcal{TC}}
\newcommand{\impactcapexgeneration}{\Phi_{n,s,t}}
\newcommand{\impactcapexflow}{\Phi_{\ell,t}}

%math 
\newcommand{\resultsin}[1]{\hspace{12pt} \bot  \hspace{12pt} #1}
\newcommand{\Forall}[1]{\hspace{20pt} \forall \,\, #1 }
\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}

\begin{document}

\title{From Linear Optimization to Transmission Cost Allocation}
\author{Fabian Hofmann}

\maketitle

% \begin{abstract}
% The abstract text goes here.
% \end{abstract}

\subsubsection*{Lagrange Mutliplier}

Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$

\begin{align}
 \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
\end{align}
...

\section*{Linear Energy Modelling and LMP}

We linearly optimize the capacity and dispatch of a simple power system.

\begin{align}
    \underset{\demand, \generation, \capacityGeneration}{\text{max}}
    \left(\utility(\demand) - \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
\end{align}

where 
\begin{itemize}
\item[] $\utility$ denotes the utility function per bus $n$, demand type $a$ time step $t$ 
 \item[] $\demand$ denotes the eletric demand 
 \item[] $\capexGeneration$ denotes the capital expenditure (CAPEX) per node $n$ and generator type $s$
 \item[] $\capacityGeneration$ denotes the generation capacity
 \item[] $\opexGeneration$ denotes the operational cost (OPEX)
 \item[] $\generation$ denotes the net generation in MW
 \item[] $\capexFlow$ denotes the CAPEX per transmission line 
 \item[] $\capacityFlow$ denotes the transmission capacity.
\end{itemize}
% 
\subsubsection*{Constraints}
In the following we neglect the utility $\utility$ of the nodal demand while fixing the demand $\demand$ to a predefined time-series. \\

The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
\begin{align}
    \sum_l \incidence \, \flow - \sum_s \generation + \sum_a \demand &= 0 \resultsin{\lmp} \Forall{n,t}
    \label[constraint]{eq:nodal_balance_lin}
\end{align}
Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay. \\

The generation $\generation$ is constraint to its nominal capacity
\begin{align}
 \generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
 \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
 - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{n,s,t} 
 \label[constraint]{eq:lower_generation_capacity_constraint}
 \end{align}
where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators. The constraints yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ which due to complementary slackness,
\begin{align}
\muuppergeneration \left( \generation - \generationpotential \, \capacityGeneration \right)  &= 0  \Forall{n,s,t} 
\label{eq:complementary_slackness_upper_generation} \\
\mulowergeneration  \, \generation &= 0 \Forall{n,s,t}
\label{eq:complementary_slackness_lower_generation} 
\end{align}
are only non-zero if the corresponding constraint is binding. \\

The transmission capacity $\capacityFlow$ limits the flow $\flow$ in both directions, such that 
\begin{align}
 \flow - \capacityFlow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
 \label[constraint]{eq:upper_flow_capacity_constraint} \\
 - \flow - \capacityFlow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
 \label[constraint]{eq:lower_flow_capacity_constraint}
\end{align}
The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the trasmission capacity in positive or negative direction, i.e. \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding. The complementary slackness 
\begin{align}
 \muupperflow \left( \flow - \capacityFlow \right)  &= 0 \Forall{\ell,t}
 \label{eq:complementary_slackness_upper_flow} \\
 \mulowerflow \left( \flow - \capacityFlow \right) &=  0 \Forall{\ell,t}
 \label{eq:complementary_slackness_lower_flow} 
\end{align}
set the respective KKT for flows staying below the thermal limit to zero. 
\\

\subsubsection*{Flow is a passive variable}
We treat the flow $\flow$ an a passive quantity only that is it is only influenced by the nodal injection.

The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit, commonly one MW, of net power production at bus $n$. Thus with a fix demand $\demand$, they direclty link the generation $\generation$ to the flow on each line according to
\begin{align}
 \flow\left( \generation\right)  = \sum_n \ptdf \left( \sum_s \generation- \sum_a \demand \right)  
 \label{eq:flow_from_ptdf}
\end{align}
The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system. It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses. The choice of slack modifies the PTDF accordingly 
\begin{align}
 \ptdf = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
 \label{eq:ptdf_slacked}
\end{align}
where $\ptdfEqual$ denote the PTDF with equally distributed slack.
When bus $n$ injects excess power, is has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack. Summing over all ingoing and outgoing flow changes resulting from an positive injection at $n$ yields again the slack 
\begin{align}
\sum_\ell \incidence[m] \, \ptdf =  \delta_{m,n} - \slack[m] 
\label{eq:slack}
\end{align}
where $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
% Sensitivity of flow $\flow$ against changes of the power production $\generation$  
% \begin{align}
%  \pdv{\flow}{\generation}  = \ptdf \label{eq:flow_sensitity} 
% \end{align}

\subsubsection*{Breaking Down the Full Lagrangian}
\begin{align}
 \lagrangian\left(\generation, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   = &- \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
 &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
%  &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \label[constraint]{eq:langrange_cycle_constraint} \\ 
 &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
 &- \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) + \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
\end{align}
% 
where $\boldsymbol{\lambda} = \left\lbrace \lmp \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables. The global maximum of the Lagrangian requires stationarity with respect to all variables. The stationarity of the generation capacity variable leads to 
\begin{align}
 \pdv{\lagrangian}{\capacityGeneration}  = 0 \hspace{10pt} \rightarrow \hspace{10pt} \capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
 \label{eq:capexGeneration_duality}
\end{align}

the stationarity of the transmission capacity to
\begin{align}
 \pdv{\lagrangian}{\capacityFlow} = 0 \hspace{10pt} \rightarrow \hspace{10pt} \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
 \label{eq:capexFlow_duality}
\end{align}

and the stationarity of the generation to 
\begin{align}
 \pdv{\lagrangian}{\generation} &= 0  &\Forall{n,s} \\
 \rightarrow \hspace{10pt}  \opexGeneration &=  \lmp - \muuppergeneration + \mulowergeneration - \sum_{\ell} \left( \muupperflow - \mulowerflow\right)  \, \ptdf - \sum_{m,\ell}\lmp[m] \, \incidence[m] \, \ptdf   &\Forall{n,s} \label{eq:opex_duality}
\end{align}
% 
For the latter, we used \cref{eq:flow_from_ptdf} which sets the derivative of the flow with respect to the generation to 
\begin{align}
\pdv{\flow}{\generation} = \ptdf                                                                                                                                                   \end{align}
\Cref{eq:capexGeneration_duality,eq:capexFlow_duality,eq:opex_duality} show how the capital and operational prices translate into dual variables. \\

Note that \cref{eq:opex_duality} must hold for every choice of slack in the PTDF. According to \cref{eq:ptdf_slacked,eq:slack}, setting the slack to $\slack = \delta_{m,n}$ results in $\ptdf = \sum_\ell \incidence[m] \, \ptdf = 0$. This leads to our first represantion for Locational Market Price, which we will refer to as the ``Island Solution``,
\begin{align}
\lmp  =  \opexGeneration + \muuppergeneration - \mulowergeneration \Forall{n,s,t}
\label{eq:lmp1}
\end{align}
Accordingly, the LMP is directly determined by the local operational price and prices for the generation capacity constraint. 
However, from the Island Solution, we can derive a second representation for the LMP. Feeding \cref{eq:lmp1} back into \cref{eq:opex_duality} and applying \cref{eq:slack}, finally leads to 
\begin{align}
\lmp =  - \sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf + \sum_m \left( \opexGeneration[m] + \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] \Forall{n,s,t} 
\label{eq:lmp2}
\end{align}
This equation generalizes the former Island Solution for the LMP which again can be reproduced by setting $\slack[m] = \delta_{m,n}$. It depicts the interdependence of the LMP, that is how $\lmp$ can be decomposed to operational prices and prices for capacity constraints from all generators and transmission lines in the system.

\subsubsection*{Cost Allocation}

The choice of slack $\slack[m]$ in \cref{eq:lmp2} decides on the generators and lines to which the nodal price at $n$ is allocated to. As we will see, there are various valid options for this choice. Mutliplied by the nodal consumption $\demandnodal = \sum_a \demand$, the slack translates into the peer-to-peer relation 
\begin{align}
 \allocatePeer = \demandnodal \, \slack[m]   \Forall{n,t}
 \label{eq:allocate_peer}
\end{align}
that is the power which is produced at bus $m$ at time $t$ and assigned to consumers at bus $n$. Or the other words, the part of demand $\demandnodal$ that is covered by generators at bus $m$. \\

Likewise, multipying the PTDF with the nodal demand will result in the flow induces by the net consumption at $n$ coming from the distributed slack 
\begin{align}
 \sum_m \allocateTransaction = - \demandnodal  \, \ptdf \Forall{n,\ell,t}  
 \label{eq:allocate_transaction}
\end{align}
where the PTDF include the slack $\slack[m]$ according to \cref{eq:ptdf_slacked}. \\

The slack $\slack[m]$ in \cref{eq:allocate_peer,eq:allocate_transaction} should be chosen such that the sum over all recipients $n$ in $\allocatePeer$ results in the nodal production, thus
\begin{align}
 \sum_n \allocatePeer = \sum_s \generation
\end{align}
Then, weighted by the nodal production share $\generationshare[m] = \generation[m]/\sum_s \generation[m]$ the allocation distributes all power pruduced by generator $(m,s)$,
\begin{align}
  \sum_n \generationshare[m] \, \allocatePeer = \generation[m] \Forall{m,s,t}
  \label{eq:recipients_sum}
\end{align}
and the sum over all producers $m$ and recipients $n$ in $\allocateTransaction$ equals the power flow on line $\ell$,
\begin{align}
 \sum_{m,n} \allocateTransaction = \flow
\end{align}

These two quantities give a more intuitive insight on how the nodal payments can be decomposed. Multiplying both sides of \Cref{eq:lmp2}  with the nodal demand $\demandnodal$ will breakdown the costs payed by consumers at node $n$ to the different expeditures

\begin{align}
 \lmp \, \demandnodal = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration
\end{align}

with the allocated payments: \\
\begin{align}
 \allocateOpex &= \sum_{m,s} \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
 &\text{OPEX for generators} 
 \label{eq:allocate_opexGeneration}\\
 \allocateCapexGeneration &= \sum_m \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
 &\text{CAPEX for the generators} 
  \label{eq:allocate_capexGeneration} \\
 \allocateCapexFlow &=  \sum_{m,\ell} \left( \muupperflow - \mulowerflow\right) \allocateTransaction  
 &\text{CAPEX for the transmission system} 
  \label{eq:allocate_capexFlow}
\end{align}

Let's have a look at the allocated OPEX first. Consumers at bus $n$ retrieve power from different generators $(m,s)$ each producing a share of the nodal production. Hence consumers at $n$ compensate the OPEX proportionally to the assigned production.

 In this way (mutliply both sides of \cref{eq:recipients_sum} by $\opexGeneration$ and sum over $t$), all OPEX payments to generator (m,s) sum up to the total OPEX spent at (m,s). \\

The CAPEX allocation for generators reveil a similar relation. According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not. If $\muuppergeneration > 0$, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding. Thus it is these times steps which push investments in $\capacityGeneration$. If in contrast $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary. 
When summing over all CAPEX payments to generator $(m,s)$ each generator retrieves exactly the cost that were spent to build the capacity $\capacityGeneration$,

\begin{align}
 \sum_{n,t} \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer = \capexGeneration \, \capacityGeneration
\end{align}
where we used \cref{eq:complementary_slackness_upper_generation,eq:capexGeneration_duality}. Throughout all time steps, each generator $(m,s)$ receives the money it spends for invesments and operation (non-profit rule). 
% The non-profit rule can also be seen when looking at the the total revenue per generator, 
% \begin{align}
%  \sum_t \lmp \, \generation &= \sum_t \opexGeneration \generation + \muuppergeneration \generation - \mulowergeneration \, \generation &\Forall{n,s} \\
%  &= \sum_t \opexGeneration \, \generation + \capexGeneration \, \capacityGeneration &\Forall{n,s}
% \end{align}
%  where we multiplied both sides of \cref{eq:lmp1} by the generation and using the complementary slackness.  
\\

The allocation of CAPEX for the transmission system $\allocateCapexFlow$ builds up on the KKT variables $\muupperflow$ and $\mulowerflow$. Again the latter translate to the necessity of transmission invesments at $\ell$ at time $t$. Consumers which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compenstations for the resulting investments. Again the sum of all CAPEX payments to line $\ell$ equal the total CAPEX spent, thus

\begin{align}
 \sum_{m,n,t} \left( \muupperflow - \mulowerflow\right) \allocateTransaction = \capexFlow \, \capacityFlow  
\end{align}
where we used the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} and the fact that summing over all sources $m$ and sinks $n$ the allocation equals the actual power flow 
\begin{align}
\sum_{n,m} \allocateTransaction = \flow
\end{align}
% A similar relation applies for the transmission CAPEX. Mulplying both side of \cref{eq:opex_duality} by the nodal injection $\sum_s \generation - \sum_a \demand$ and using the complementary slackness ...
% shows that the CAPEX per line amounts to the total congestion revenue
% \begin{align}
% \capexFlow \, \capacityFlow &= - \sum_{n,t} \lmp \, \incidence \, \flow \Forall{\ell} \label{eq:non_profit_branch}
% \end{align}
% The transmission lines carry power from nodes with lower LMP to nodes with higher LMP, the price difference times the transported power is their congestion revenue. \note{Sign is different to Tom's paper (there a minus is missing).}

\subsubsection*{Adding CO$_2$ Constraints}

Imposing an additional CO$_2$ constraint limiting the total emission to K,  
\begin{align}
 \sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionPrice} 
 \label[constraint]{eq:co2_constraint}
\end{align}
with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionPrice$ in \euro/tonne-CO$_2$. 
% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
% 
% \begin{align}
% \capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
% \label{eq:non_profit_generator_emission}
% \end{align}
% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
\begin{align}
\opexGeneration \rightarrow \opexGeneration + \emission \, \emissionPrice \label[relation]{eq:shift_opex_by_emission_cost}
\end{align}
This leads to allocated CO$_2$ cost compensation of node $n$ of
 \begin{align}
 \allocateEmissionCost &= \emissionPrice \, \sum_{m,s} \emission[m] \, \generationshare[m] \, \allocatePeer \Forall{n,t} \label{eq:allocate_emissionPrice}
\end{align}

% ----------------------------------------------------------------------------------------
% ----------------------------------------------------------------------------------------

% The power injection per node and time denotes the nodal power imbalance between generation and demand
% 
% \begin{align}
%  \injection = \sum_s \generation - \sum_a \demand
% \end{align}
% The net production is the amount of positive injection per node and time 
% \begin{align}
%  \netproduction = \text{min}\left( \injection, 0 \right) 
% \end{align}
% The nodal net consumption denotes the amount of negative injection 
% \begin{align}
%  \netconsumption = \text{min}\left( - \injection, 0 \right) 
% \end{align}
% The nodal self-consumption equals the minimum of the nodal production and demand. 
% \begin{align}
%  \selfconsumption = \text{min}\left( \netproduction, \netconsumption \right) 
% \end{align}

\section*{Showcase}

\subsection*{Network with CO$_2$ constraint}
We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}. It consists of nine buses and ten time steps. The solver optimizes the capacity of two generators, wind and gas, per bus. ...

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{compare_allocation.png}
    \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ constraint. The left bars consist of the allocated OPEX $\allocateOpex$, the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$, while the right bars show the of the nodal consumption times the LMP. }
    \label{fig:cost_allocation}
\end{figure}

\begin{figure}[h]
\begin{subfigure}{.5\textwidth}
\centering
 \includegraphics[width=\textwidth]{network.png}
 \caption{}
 \label{fig:network}
\end{subfigure}
\begin{subfigure}{.5\textwidth}   
    \centering
    \includegraphics[width=\textwidth]{nodal_payments.png}
    \caption{}
    \label{fig:nodal_payments}
\end{subfigure}
\caption{Network used for showcasing. (a) shows the distributing of generation capacities $\capacityGeneration$, the widths of the transmission lines are proportional to their thermal limit $\capacityFlow$. (b) shows the total nodal payments according to the cost allocation.}
\end{figure}

% 
% 
% 
% \newpage
% \subsubsection*{Relaxed CO$_2$ Constraint}
% 
% 
% \begin{figure}[h]
% \begin{subfigure}{.5\textwidth}
% \centering
%  \includegraphics[width=\textwidth]{network_relaxed_co2.png}
%  \caption{}
%  \label{fig:network_relaxed_co2}
% \end{subfigure}
% \begin{subfigure}{.5\textwidth}   
%     \centering
%     \includegraphics[width=\textwidth]{nodal_payments_relaxed_co2.png}
%     \caption{}
%     \label{fig:nodal_payments_relaxed_co2}
% \end{subfigure}
% \caption{Similar to \cref{fig:network} and \cref{fig:nodal_payments} but without CO$_2$ constraint.}
% \end{figure}
% 
% \begin{figure}[h]
%     \centering
%     \includegraphics[width=\textwidth]{compare_allocation_relaxed_co2.png}
%     \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ \cref{eq:co2_constraint}. Only one time-step $t=6$ determines the allocation of generator CAPEX $\allocateCapexGeneration$, as for all other time-steps \cref{eq:upper_generation_capacity_constraint} is not binding. Again note the cost scale difference between time step 6 and all others.}
%     \label{fig:cost_allocation_relaxed_co2}
% \end{figure}
% 
% \subsection*{How does the cost flow through the network}
% 
% \begin{figure}[h]
%     \begin{subfigure}{.5\textwidth}
%       \centering
%       \includegraphics[width=\textwidth]{opex_flow.png}
%       \label{fig:opex_flow}
%     \end{subfigure}%
%     \begin{subfigure}{.5\textwidth}
%       \centering
%       \includegraphics[width=\textwidth]{capex_flow.png}
%       \label{fig:capex_flow}
%     \end{subfigure}
%     \caption{}
%     \label{fig:fig}
% \end{figure}

% \begin{appendix}
% Since the P2P comprises all net production and net consumption summing over all sources must result in the nodal demand  
% 
% \begin{align}
%  \sum_{n} \allocatePeer = \sum_a \demand[m]
% \end{align}
% as well as summing over all sinks results in the nodal generation
% \begin{align}
%  \sum_{m} \allocatePeer = \sum_s \generation
% \end{align}
% 
% The P2P relation can be further broke down to allocating generator specific $ \generationshare \, \allocatePeer $ where $\generationshare = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodel production. Similarly the allocation to consumers a are given by and consumer specific $ \demandshare[m] \allocatePeer$ with the nodal comsumer share given by $\demandshare = \demand/\sum_a \demand$.
% \end{appendix}

% \bibliographystyle{ieeetr}
% \printbibliography

\end{document}