### 2020-06-27 paper/playaround.tex
play around with tom's approach

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{amsmath}
\usepackage[colorlinks]{hyperref}
\usepackage[backend=biber,style=draft]{biblatex}
\usepackage{eurosym}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{enumitem} % for alphabetical enumeration 
\usepackage{accents}
\usepackage[capitalise]{cleveref}

%opening
\title{}
\author{Fabian Hofmann}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

% style operators
\newcommand{\ie}{\textit{i.e.} }
\newcommand{\eg}{\textit{e.g.} }
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\newcommand{\note}[1]{\textcolor{Orange}{#1}}
\newcommand{\vpad}{\vspace{1mm}}
\newcommand{\hpad}{\hspace{15pt}}

% general symbols
\newcommand{\generation}[1][n]{g_{#1,s,t}}
\newcommand{\generationpotential}{\bar{g}_{n,s,t}}
\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
\newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
\newcommand{\capacityGeneration}{G_{n,s}}
\newcommand{\capacityGenerationUpper}{\bar{G}_{n,s}}
\newcommand{\capacityGenerationLower}{\ubar{G}_{n,s}}
\newcommand{\capacityFlow}{F_{\ell}}
\newcommand{\capacityFlowUpper}{\bar{F}_{\ell}}
\newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
\newcommand{\capexGeneration}{c_{n,s}}
\newcommand{\capexFlow}{c_{\ell}}
\newcommand{\opexGeneration}[1][n]{o_{#1,s}}
\newcommand{\demand}[1][n]{d_{#1,a,t}}
\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
\newcommand{\utility}{U_{n,a,t}}
\newcommand{\incidence}[1][n]{K_{#1,\ell}}
\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
\newcommand{\ptdfEqual}[1][n]{\ptdf[#1]^\circ}
\newcommand{\slackflow}{k_{\ell}}
\newcommand{\slack}[1][n]{k_{#1}}
\newcommand{\slackk}[1][n]{k^*_{#1}}
\newcommand{\Slack}{k_{m,n}}
\newcommand{\Slackk}{k^*_{m,n}}
\newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
\newcommand{\muuppergeneration}[1][n]{\bar{\mu}_{#1,s,t}}
\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
\newcommand{\flow}{f_{\ell,t}}
\newcommand{\cycle}{C_{\ell,c}}
\newcommand{\impedance}{x_\ell}
\newcommand{\cycleprice}{\lambda_{c,t}}
\newcommand{\injection}{p_{n,t}}
\newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
\newcommand{\netproduction}[1][n]{p^{+}_{#1,t}}
\newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}

% matrix notation
\newcommand{\incidenceM}{K}
\newcommand{\flowM}{f}
\newcommand{\injectionM}{p}
\newcommand{\slackM}{k}
\newcommand{\DirectedIncidence}{\mathcal{K}}
\newcommand{\InverseAPInjection}{\mathcal{J}}
\newcommand\diag[1]{\operatorname{diag}\left(#1\right)}

% totals
\newcommand{\totalnetconsumption}{p^{-}_{t}}
\newcommand{\totalnetproduction}{p^{+}_{t}}
\newcommand{\totalselfconsumption}{p^{\circ}_{t}}

\newcommand{\lagrangian}{\mathcal{L}}

% allocation quantities
\newcommand{\allocatePeer}[1][(m, s) \rightarrow n]{A_{#1,t}}
\newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
\newcommand{\allocateTransaction}[1][m \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}

\newcommand{\emission}[1][n]{e_{#1,s}}
\newcommand{\emissionPrice}{\mu_{\text{CO2}}}
\newcommand{\megawatthour}{MWh$_\text{el}$}
\newcommand{\totalcost}{\mathcal{TC}}
\newcommand{\totalOpexGeneration}{\mathcal{O}}
\newcommand{\totalOpexFlow}{\mathcal{O}^F}
\newcommand{\totalCapexGeneration}{\mathcal{C}^G}
\newcommand{\totalCapexFlow}{\mathcal{C}^F}
\newcommand{\totalEmissionCost}{\mathcal{E}}

%math 
\newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
\newcommand{\pdv}[2]{\dfrac{\partial #1}{\partial #2}}

The total cost $\totalcost$ are composed of
\begin{align}
\totalcost &= \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow  \\
&= \sum_{n,t} \left(  \pdv{\totalOpexGeneration}{\nodaldemand} + \pdv{\totalCapexGeneration}{\nodaldemand} + \pdv{\totalCapexFlow}{\nodaldemand} \right) \nodaldemand 
\label{eq:total_cost_terms}
\end{align}
% 
The total OPEX is given by 
\begin{align}
 \totalOpexGeneration& = \sum_{m,s,t} \opexGeneration[m]\,  \generation[m] 
\end{align}
% 
The OPEX payments assigned to node $n$ at time step $t$ is given by 
% 
\begin{align}
\allocateOpex  = \pdv{\totalOpexGeneration}{\nodaldemand} \, \nodaldemand & =  \sum_{m,s,t}\opexGeneration[m] \,  \allocatePeer
\end{align}
% 
with $\allocatePeer$ being defined as the power produced by generator $(m, s)$ and consumed at node $n$
\begin{align}
 \allocatePeer = \pdv{\generation[m]}{\nodaldemand} \, \nodaldemand
\end{align}
% 
The derivative $\partial \generation[m] / \partial \nodaldemand$ is not well-defined but implies a degree of freedom. However it is constraint to allocating exactly the amount of power which is produced by generator $(m,s)$, thus 
\begin{align}
\generation[m] = \sum_n \allocatePeer 
\label{eq:solution_space_slack}
\end{align}
% 
The direct payment of node $n$ to generator $(m,s)$ is then  
% 
\begin{align}
 \allocateOpex[n \rightarrow (m,s)] &= 
\opexGeneration[m] \,  \pdv{\generation[m]}{\nodaldemand} \, \nodaldemand
\label{eq:allocate_opexGeneration_detailed}\\
\end{align}

% \begin{align}
% \allocateCapexGeneration[n \rightarrow (m,s)] &= 
% \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
% \label{eq:allocate_capexGeneration_detailed} \\
% \allocateCapexFlow[n \rightarrow \ell] &=  
% \left( \muupperflow - \mulowerflow\right) \sum_{m} \allocateTransaction  
% \label{eq:allocate_capexFlow_detailed}
% \end{align}

The CAPEX for the transmission system is given by

\begin{align}
 \totalCapexFlow & = \sum_\ell \capexFlow \capacityFlow = \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right)  \flow 
\end{align}

The individual cost terms are 
\begin{align}
\allocateCapexFlow = \pdv{\totalCapexFlow}{\nodaldemand}\nodaldemand & = \pdv{\left( \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right)  \flow \right) }{\nodaldemand} \\
&= \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \ptdf 
\end{align}

\begin{align}
 \pdv{\muupperflow}{\nodaldemand} = 0 \;\; ?
\end{align}

\begin{align}
\pdv{\lagrangian}{\demand} & = \pdv{\lagrangian}{\muupperflow} \,  \pdv{\muupperflow} {\nodaldemand} \\
&= \left( \flow - \capacityFlow \right)  \,  \pdv{\muupperflow} {\nodaldemand}
\end{align}

\begin{align}
\notag
 \pdv{\lagrangian}{\demand}
 =&    \sum_{m, s, t} \opexGeneration[m] \Slack  \\
\notag
&+ \sum_{m,t} \lmp[m] \left(\sum_\ell \incidence[m] \, \ptdf  - \Slack + \delta_{m,n}  \right)  \\ 
\notag
&+ \sum_{n,s,t} \muuppergeneration \generationshare \Slack %  - \sum_{n,s,t} \mulowergeneration \generation  
\\
&+ \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \ptdf      
\label{eq:full_lagrangian}
\end{align}

\subsubsection*{Production CAPEX}
\begin{align}
 \totalCapexGeneration& = \sum_{n,s} \capexGeneration \capacityGeneration\\
 &= \sum_{n,t} \muuppergeneration  \generation 
\end{align}

and taking the derivative

\begin{align}
\pdv{\totalCapexGeneration}{\nodaldemand} & = \pdv{ \left(  \sum_{m,t} \muuppergeneration[m]  \generation[m] \right) }{\nodaldemand} \\
&= \sum_{m,t} \generationshare  \muuppergeneration \, \Slack
\end{align}

\end{document}