### 2020-10-13 paper/flow_based.tex
write intro text (without citations); now change text for a better story

\documentclass[11pt,twocolumn]{article}
\usepackage{graphicx}
\usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{amsmath,bm}
\usepackage[colorlinks]{hyperref}
\usepackage[backend=biber, bibencoding=utf8]{biblatex}
\usepackage{eurosym}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{booktabs} % for table (toprule, bottomrule)
\usepackage{enumitem} % for alphabetical enumeration 
\usepackage{accents}
\usepackage[capitalise]{cleveref}

\addbibresource{main.bib}
\graphicspath{{../figures/}{../figures/example/}}

\crefname{relation}{Rel.}{Rels.}
\creflabelformat{relation}{(#2#1#3)}
\crefname{constraint}{Constr.}{Constrs.}
\creflabelformat{constraint}{(#2#1#3)}
\crefname{subequations}{Eqs.}{Eqs.}
\creflabelformat{subequations}{(#2#1#3)}

\setlength\parindent{8pt}

% style operators
\newcommand{\ie}{\textit{i.e.} }
\newcommand{\eg}{\textit{e.g.} }
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\newcommand{\note}[1]{\textcolor{Orange}{#1}}
\newcommand{\vpad}{\vspace{1mm}}
\newcommand{\hpad}{\hspace{15pt}}
\newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}

% general symbols
\newcommand{\state}{s_{i,t}}
\newcommand{\capacity}{S_{i}}
\newcommand{\costfactor}{\gamma_{i,t}}
\newcommand{\capacityupper}{\bar{S}}
\newcommand{\capacitylower}{\ubar{S}}
\newcommand{\muuppernom}{\bar{\mu}^\text{nom}_{i}}
\newcommand{\mulowernom}{\ubar{\mu}^\text{nom}_{i}}

\newcommand{\kk}{k~\euro\,}

%generation
\newcommand{\generation}{g_{s,t}}
\newcommand{\generationpotential}{\bar{g}_{s,t}}
\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
\newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
\newcommand{\capacitygeneration}{G_{s}}
\newcommand{\capacitygenerationupper}{\bar{G}_{s}}
\newcommand{\capacitygenerationlower}{\ubar{G}_{s}}
\newcommand{\operationalpricegeneration}{o_{s}}
\newcommand{\capitalpricegeneration}{c_{s}}
\newcommand{\mulowergeneration}{\ubar{\mu}_{s,t}}
\newcommand{\muuppergeneration}{\bar{\mu}_{s,t}}
\newcommand{\muuppergenerationnom}{\bar{\mu}^\text{nom}_{s}}
\newcommand{\mulowergenerationnom}{\ubar{\mu}^\text{nom}_{s}}

% flow
\newcommand{\flow}{f_{\ell,t}}
\newcommand{\capacityflow}{F_{\ell}}
\newcommand{\capacityflowUpper}{\bar{F}_{\ell}}
\newcommand{\capacityflowLower}{\ubar{F}_{\ell}}
\newcommand{\operationalpriceflow}{o_\ell}
\newcommand{\capitalpriceflow}{c_{\ell}}
\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
\newcommand{\muupperflownom}{\bar{\mu}^\text{nom}_{\ell}}
\newcommand{\mulowerflownom}{\ubar{\mu}^\text{nom}_{\ell}}

% storage
\newcommand{\storage}{g_{r,t}}
\newcommand{\storagedispatch}{\storage^\text{dis}}
\newcommand{\storagecharge}{\storage^\text{sto}}
\newcommand{\storagesoc}{\storage^\text{ene}}
\newcommand{\storageprevioussoc}{g_{r,t-1}^\text{ene}}

\newcommand{\efficiency}{\eta_{r}}
\newcommand{\efficiencydispatch}{\efficiency^\text{dis}}
\newcommand{\efficiencycharge}{\efficiency^\text{sto}}
\newcommand{\efficiencysoc}{\efficiency^\text{ene}}

\newcommand{\operationalpricestorage}{o_r}
\newcommand{\capitalpricestorage}{c_r}
\newcommand{\capacitystorage}{G_r}
\newcommand{\capacitystorageupper}{\bar{G}_r}
\newcommand{\capacitystoragelower}{\ubar{G}_r}
\newcommand{\mulowerstorage}{\ubar{\mu}_{r,t}}
\newcommand{\muupperstorage}{\bar{\mu}_{r,t}}
\newcommand{\mulowerstoragedispatch}{\ubar{\mu}_{r,t}^\text{dis}}
\newcommand{\muupperstoragedispatch}{\bar{\mu}_{r,t}^\text{dis}}
\newcommand{\mulowerstoragecharge}{\ubar{\mu}_{r,t}^\text{sto}}
\newcommand{\muupperstoragecharge}{\bar{\mu}_{r,t}^\text{sto}}
\newcommand{\mulowerstoragesoc}{\ubar{\mu}_{r,t}^\text{ene}}
\newcommand{\muupperstoragesoc}{\bar{\mu}_{r,t}^\text{ene}}
\newcommand{\muupperstoragenom}{\bar{\mu}^\text{nom}_r}
\newcommand{\mulowerstoragenom}{\ubar{\mu}^\text{nom}_r}
\newcommand{\mustateofcharge}{\lambda^\text{ene}_{r,t}}
\newcommand{\munextstateofcharge}{\lambda^\text{ene}_{r,t+1}}

% other
\newcommand{\lagrangian}{\mathcal{L}}
\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
\newcommand{\averagelmp}{\left<\lmp\right>_t}
\newcommand{\demand}[1][n]{d_{#1,t}}
\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
\newcommand{\injection}[1][n]{p_{#1,t}}
\newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
\newcommand{\netproduction}[1][n]{p^{+}_{#1,t}}
\newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}
\newcommand{\totalnetconsumption}{p^{-}_{t}}
\newcommand{\totalnetproduction}{p^{+}_{t}}
\newcommand{\totalselfconsumption}{p^{\circ}_{t}}

% network
\newcommand{\incidence}[1][n]{K_{#1,\ell}}
\newcommand{\incidencegenerator}[1][n]{K_{#1,s}}
\newcommand{\incidencestorage}[1][n]{K_{#1,r}}
\newcommand{\incidenceasset}[1][n]{K_{#1,i}}
\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
\newcommand{\cycle}{C_{\ell,c}}
\newcommand{\reactance}{x_\ell}
\newcommand{\cycleprice}{\lambda_{c,t}}

% Cost Terms
\newcommand{\emission}{e_{s}}
\newcommand{\emissionprice}{\mu_{\text{CO2}}}
\newcommand{\megawatthour}{MWh$_\text{el}$}
\newcommand{\totalcost}{\mathcal{T}}
\newcommand{\cost}{\mathcal{C}}
\newcommand{\opex}{\mathcal{O}}
\newcommand{\opexgeneration}{\mathcal{O}^G}
\newcommand{\opexflow}{\mathcal{O}^F}
\newcommand{\opexstorage}{\mathcal{O}^E}
\newcommand{\capexgeneration}{\mathcal{C}^G}
\newcommand{\capexflow}{\mathcal{C}^F}
\newcommand{\capexstorage}{\mathcal{C}^E}
\newcommand{\emissioncost}{\mathcal{E}}
\newcommand{\remainingcost}{\mathcal{R}}
\newcommand{\scarcitycost}{\remainingcost^\text{scarcity}}
\newcommand{\subsidycost}{\remainingcost^\text{subsidy}}

% allocation quantities
\newcommand{\allocategeneration}[1][s, n]{A_{#1,t}}
\newcommand{\allocatestoragedispatch}[1][r, n]{A_{#1,t}}
\newcommand{\allocatepeer}[1][m \rightarrow n]{A_{#1,t}}
\newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
\newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocatestate}[1][i, n]{A_{#1,t}}

\newcommand{\allocatecost}[1][n \rightarrow i]{\cost_{#1, t}}
\newcommand{\allocatecapexgeneration}[1][n \rightarrow s]{\capexgeneration_{#1,t}}
\newcommand{\allocatecapexflow}[1][n \rightarrow \ell]{\capexflow_{#1,t}}
\newcommand{\allocatecapexstorage}[1][n \rightarrow r]{\capexstorage_{#1,t}}
\newcommand{\allocateopex}[1][n \rightarrow s]{\opex_{#1,t}}
\newcommand{\allocateemissioncost}[1][n \rightarrow s]{\emissioncost_{#1,t}}
\newcommand{\allocatescarcitycost}[1][n \rightarrow i]{\scarcitycost_{#1,t}}
\newcommand{\allocatesubsidycost}[1][n \rightarrow i]{\subsidycost_{#1,t}}

\begin{document}

\title{Tracing prices: A flow-based cost allocation for optimized power systems}
\author{F. Hofmann}

\maketitle

\begin{abstract}
% Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling. It leads to perfectly scheduled operations of generators and ... 
\end{abstract}

\subsection*{Highlights}
\begin{itemize}
 \item Flow Allocation schemes determine peer-to-peer (P2P) allocation between sources and sinks 
 \item P2P allocation can be used to assign payments of consumers to single network assets
 \item Using shadow prices, Capital Expenditures are accurately allolcated to users which profit from investments
 \item The P2P cost assignments are in alignment with the nodal pricing scheme of a welfare-maximized system   
 \item This opens the possibility to transparently integrate transmission constraints and investment costs in national and international development plans 
 \item Applying to a German future scenario shows how higher electricity costs arise for buses which rely on investments in transmission and local backup generation  
\end{itemize}

\subsubsection*{Nomenclature}

\begin{table}[h]
    \centering
    \begin{tabular}{ll}
        $\lmp$ & Locational Market Price at bus $n$ and  \\ & time step $t$ in \euro/MW \vpad \\
        $\demand$ & Electric demand per bus $n$, demand type $a$, \\ & time step $t$ in MW  \vpad \\
        $\generation$ & Electric generation of generator $s$, \\ & time step $t$  in MW \vpad \\
        $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
        $\operationalpricegeneration$ & Operational price in \euro/MW \vpad \\
        $\capitalpricegeneration$ & Capital Price in \euro/MW \vpad \\
        $\capitalpriceflow$  & Capital Price in \euro/MW for transmission \\ & capacity on line $\ell$  \vpad \\
        $\capacitygeneration$ & Generation capacity in MW \vpad \\
        $\capacityflow$ & Transmission capacity in MW \vpad \\
        $\incidence$ & Incidence matrix \vpad 
    \end{tabular}
\end{table}

\section{Introduction}

Today's power systems are subject to a deep and ongoing transformation. The needed shift to from controllable to weather-driven power generation, the constant improvement and innovation of technology require rigorous system planning and international cooperation. The core of the problem manifests in the costs. Firstly, these should be as small as possible while meeting ecological and sociological standards. Secondly, they must be distributed in a fair and transparent manner. Cost-driving aspects or market players should be detected and addressed appropriately. 
In this context, power system modelling has increasingly gained attention during the last years. While many studies show how cost-efficiency and renewable energy are brought together, the question of how and, more importantly, upon what grounds costs are distributed was often left open. \\
This study fills this gap. On the basis of an optimized network the operational state of each time step and corresponding characteristics are considered to allocate all system costs. Therefore, two basic concepts are used: Firstly, the revenue of an network asset at the cost-optimum exactly matches its operational and capital expenditures, also known as the zero-profit condition. Secondly, \textit{flow tracing}, following Bialek's Average Participation, assigns the operation of single network assets to consumers in an plausible manner. 
The following shows how combining these allow for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the end consumers.

\section{Economic Context}

In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms. Typically, these include OPEX for generators $\opexgeneration$, expenditures for emissions $\emissioncost$, CAPEX for generators $\capexgeneration$, CAPEX for the transmission system $\capexflow$ and so on, \ie
\begin{align}
\totalcost &= \opexgeneration + \emissioncost +  \capexgeneration + ...
\label{eq:total_cost}
\end{align}
In turn, each of these terms $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, ...\}$ consists of cost associated to the asset $i$ in the system, 
\begin{align}
    \cost = \sum_{i} \cost_{i}
    \label{eq:asset_cost}
\end{align} 
where an asset describes any operating part of the network, such as a generator, line, energy storage etc. %For example the total OPEX, $\opexgeneration = \sum_{s} \opexgeneration_{s}$, is the combined OPEX of all generators $s$.

In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$. Hence, it is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
\begin{align}
\lmp = \pdv{\,\totalcost}{\,\demand}
\label{eq:lmp}
\end{align}
% 
\begin{figure}[h]
\centering
\includegraphics[width=.8\linewidth]{price_decomposition.png}
\caption{Schematic decomposition of the Locational Market Price $\lmp$. In power system model with optimal long-term operation and planning, the total system costs $\totalcost$ split into different cost terms, \ie OPEX and CAPEX for production and transmission and possibly other expenditures. }
\label{fig:price_decomposition}
\end{figure}
From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into contribution to the above mentioned cost terms. This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}. 
% 
This leads to a nodal pricing where over the span of optimized time steps $t$, the system costs are partially or totally payed back by the consumers 
\begin{align}
\totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
\label{eq:total_revenue}
\end{align}
depending on the costs $\remainingcost$ which are independent of the nodal demand  
\begin{align}
 \pdv{\remainingcost}{\demand} = 0
\end{align}
% 
Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional request to the network design, such as capacity expansion limits or minimum share of one technology in the power mix. However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.

However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered. This work aims at presenting and illustrating a an intuitive, peer-to-peer cost allocation including all network assets.

\section{Dispatch-Based Cost Allocation}
\label{sec:theory}
\begin{subequations}\label[subequations]{eq:general_scheme}

Let $\cost_{i}$ denote an arbitrary cost term associated with asset $i$. Consider a long-term equilibrium in a power system with perfect competition, then, according to the zero-profit condition, each cost term $\cost_{i}$ can be considered as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$, \ie
\begin{align}
    \cost_{i} = \sum_t  \costfactor \, \state
    \label{eq:cost_decomposition}
\end{align}
where $\costfactor$ denotes a cost factor in \euro/MW. 
If $\cost_i$ describes the OPEX occasioned by asset $i$, the cost factor $\costfactor$ is simply given by the marginal operational price $o_i$. However, as we will show later, if it describes the CAPEX of asset $i$, $\costfactor$ is a composition of shadow prices $\mu_{i,t}$ given at the optimum. %As we will show later the composition of $\mu_{i,t}$ must be determined for each type of asset individually.

% where $\costfactor$ is either a fixed operational price or a composition of shadow prices
Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
\begin{align}
    \allocatecost = \costfactor \,\pdv{\state}{\demand} \demand
    \label{eq:cost_allocation}
\end{align}
The derivative on the right hand side is defined through the sensitivity of the operational variable $\state$ at the optimum against changes in the demand. However, we replace this quantity by 
\begin{align}
    % \pdv{\state}{\demand} \demand \rightarrow \int \pdv{\state}{\demand}  \,\mathrm{d}\demand  \rightarrow \allocatestate
    \pdv{\state}{\demand} \demand \Leftrightarrow  \allocatestate
    \label{eq:state_allocation}
\end{align}    
where $\allocatestate$ represents the amount of power that  asset $i$ supplies demand $\demand$ with. As we will show later, it is calculated from established flow allocation schemes. By following this approach, the differential quantity on the left side, is replaced by an integral quantity. The latter does not only rely on a marginal sensitivity of the nodal demand $\demand$, but makes further assumptions on how the operational state $\state$, all in all, serves consumers in the network. The dispatch allocation $\allocatestate$ ensures that all power produced or processed by asset $i$ at time $t$  is assigned to consumers,
\begin{align}
    \state = \sum_n \allocatestate,
    \label{eq:state_allocation_constraint}
\end{align}
from which follows that the sum of all cost contributions return the total cost term, 
\begin{align}
    \cost_i = \sum_{n,t} \allocatecost
    \label{eq:cost_payback}
\end{align}
% Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
% \begin{align}
%     \cost_{n,t} = \sum_i \allocatecost
% \end{align}

\end{subequations}

By combining \cref{eq:cost_allocation,eq:state_allocation}, we are able to formulate a full cost allocation from demand to network assets for every time-step $t$. In particular, the definition of cost factors $\costfactor$ must be derived for each cost term individually. 
In the following we depict the presented scheme for typical classes of assets. Therefore, we assume a network with generators $s$, transmission lines $\ell$ and storage units $r$, each asset $i = \{s, \ell, r\}$ adds a term for OPEX and a term for CAPEX to the total system cost $\totalcost$.

\begin{table*}[t]
    \begin{center}
        \begin{tabular}{l|c|c|c|c|c}
        & $i$ & $\cost$ & $\cost_i$  & $\costfactor$ & $\state$  \\
        \toprule 
        OPEX Production & $s$ & $\opexgeneration$ & $\sum_{t} \operationalpricegeneration \, \generation$   & $\operationalpricegeneration$ & $\generation$ \\  
        OPEX Transmission  & $\ell$ & $\opexflow$ & $\sum_{t} \operationalpriceflow \, |\flow|  $ & $\operationalpriceflow$ & $|\flow|$ \\  
        OPEX Storage & $r$  & $\opexstorage$ & $\sum_{t} \operationalpricestorage \, \storagedispatch$ &  $\operationalpricestorage$ & $\storage$ \\
        \midrule
        Emission Cost & $s$ & $\emissioncost$ & $ \emissionprice \, \emission \, \generation$ & $\emissionprice \,\emission$ & $\generation$ \\
        \midrule   
        CAPEX Production & $s$ & $\capexgeneration$ & $ \capitalpricegeneration \capacitygeneration$ & $\muuppergeneration$ & $\generation$ \\
        CAPEX Transmission & $\ell$ & $\capexflow$ & $ \capitalpriceflow \capacityflow$ & $\left(\muupperflow - \mulowerflow \right)$ & $\flow$ \\
        CAPEX Storage & $r$ & $\capexstorage$ & $ \capitalpricestorage \capacitystorage$ & $ \muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge $ & $\storage$ \\
        % \bottomrule   
    \end{tabular}
    \end{center}
    \caption{Mapping of different cost terms to the cost allocation scheme given in \cref{eq:general_scheme}. These include OPEX \& CAPEX for production, transmission and storage assets in the network, as well as a cost term for the total Green House Gas (GHG) emissions.}
    \label{tab:cost_allocation_map}
\end{table*}

\subsection{OPEX Allocation}

Let the operational price for an asset $i$ be given by $o_i$. Then, the OPEX occasioned, for example, by generator $s$ is given by 
\begin{align}
    \opexgeneration_s = \sum_t \operationalpricegeneration \, \generation 
    \label{eq:opexgeneration}
\end{align}
where $\generation$ denotes its power generation at time $t$. \Cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme}. As a result we obtain 
\begin{align}
    \allocateopex[n \rightarrow s] &= 
   \operationalpricegeneration \,  \allocategeneration
\label{eq:allocate_opexGeneration_detailed}
\end{align}
which is the contribution of $\demand$ to the OPEX at generator $s$.
The quantity  
\begin{align}
 \allocategeneration \Leftrightarrow \pdv{\generation}{\demand} \, \demand
 \label{eq:allocate_peer}
\end{align}
superseding the derivative on the right side, describes the power that is produced by generator $s$ and consumed at node $n$ at time $t$. \\
In the same manner, we can follow the scheme to allocate OPEX for flow $\flow$ and storage unit dispatch $\storagedispatch$. As we assume a bidirectional flow on line $\ell$, the OPEX is set proportional to the absolute value of the flow. The upper section in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation. \\

The scheme works for all other cost attached to the operational state of an asset $i$. Given for example a fix price for emissions $\emissionprice$ in \euro\, per tonne-CO$_2$ equivalents, the cost term for emission adds up to 
\begin{align}
 \emissioncost = \emissionprice \, \sum_s  \emission \, \generation
\end{align}
where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour\, of generator $s$.
The allocated payment for consumers at bus $n$ at time $t$ assigned to generator $s$ is then given by 
\begin{align}
 \allocateemissioncost = \emissionprice \, \emission \, \allocategeneration
\end{align}

\subsection{CAPEX Allocation}
For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and its binding constraints. For all assets, let the capital price for one MW capacity expansion be denoted by $c_i$.  All quantities for the CAPEX allocation, which we now discuss in detail, are summarized in the lower section of \cref{tab:cost_allocation_map}.

\subsubsection{Generators}

The nominal capacity $\capacitygeneration$ constrains the generation $\generation$ in the form of 
\begin{align}
\generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
\label[constraint]{eq:upper_generation_capacity_constraint}\\ 
- \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
\label[constraint]{eq:lower_generation_capacity_constraint}
\end{align}
where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators. At a cost-optimum, these two constraints yield the shadow prices $\muuppergeneration$ and $\mulowergeneration$.  As shown in \cite{brown_decreasing_2020} and in detailed in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$, 
\begin{align}
 \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
 \label{eq:no_profit_capex_generation}
\end{align}
This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition}. The resulting the CAPEX allocation is given by
\begin{align}
 \allocatecapexgeneration = \muuppergeneration \, \allocategeneration
 \label{eq:allocate_capexGeneration_detailed}
\end{align}
How does this allocation behave? According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not. If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding, which pushes investments in $\capacitygeneration$.  If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary. 
When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} , we can use \cref{eq:state_allocation_constraint,eq:no_profit_capex_generation} to see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.

\subsubsection{Transmission Lines}

The transmission capacity $\capacityflow$ limits the flow $\flow$ in both directions,
\begin{align}
\flow - \capacityflow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
\label[constraint]{eq:upper_flow_capacity_constraint} \\
- \flow - \capacityflow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
\label[constraint]{eq:lower_flow_capacity_constraint}
\end{align}
which yield the shadow prices $\muupperflow$ and $\mulowerflow$. Again, we use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which shows that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow 
\begin{align}
\capexflow_\ell = \capitalpriceflow \capacityflow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
\label{eq:no_profit_capex_flow}
\end{align}
From here, we follow the scheme in \cref{eq:general_scheme} which finally defines the CAPEX allocation as 
\begin{align}
    \allocatecapexflow &=  
   \left( \muupperflow - \mulowerflow\right) \, \allocateflow
   \label{eq:allocate_capexFlow_detailed}
\end{align}
The quantity 
\begin{align}
 \allocateflow \Leftrightarrow  \pdv{\flow}{\demand}\, \demand
\end{align}
represents the flow that the demand at node $n$ and time $t$ causes on line $\ell$.
% According to \cref{eq:state_allocation_constraint} it has to fulfill 
% \begin{align}
%  \flow = \sum_{n} \allocateflow
% \label{eq:allocate_flow_constraint}
% \end{align}
The  shadow prices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$. Hence, the definition of $\allocatecapexflow$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$. Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent. This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:state_allocation_constraint,eq:no_profit_capex_flow}

\subsubsection{Storages}
\label{sec:storages}

% The complementary slackness of the above constraints,  
% \begin{align}
%     \capexstorage_r = \capitalpricestorage \, \capacitystorage = \sum_t \muupperstoragedispatch \, \storagedispatch + \muupperstoragecharge \, \storagecharge + \muupperstoragesoc \, \storagesoc
% \end{align}

In a simplified storage model, $\capacitystorage$ limits the storage dispatch $\storagedispatch$ and charging $\storagecharge$. Further it limits the maximal storage capacity $\storagesoc$ by a fix ratio $h_r$, denoting the maximum hours at full discharge. The storage $r$ dispatches power with efficiency $\efficiencydispatch$, charges power with efficiency $\efficiencycharge$ and preserves power from one time step $t$ to the next, $t+1$, with an efficiency of $\efficiencysoc$. In \cref{sec:zero_profit_storage_units} we formulate the mathematical details. As already shown in \cite{brown_decreasing_2020}, the total expenditures at $r$ are fully paid back by differences of the LMP at which the storage ``buys`` and ``sells`` power. Taking only the CAPEX into account the zero profit condition reduces to
\begin{align}
    \notag
    \capexstorage =& \capitalpricestorage \, \capacitystorage \\
    \notag
    =& \sum_t \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \storagedispatch \\
    &- \sum_t \lmp \incidencestorage  \storagecharge \Forall{r} 
    \label{eq:no_profit_capex_storage}
\end{align}
where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower dispatch capacity bound and $\mustateofcharge$ is the shadow price of the energy balance constraint. When applying the cost allocation scheme \cref{eq:general_scheme}, it stands to reason to assume that $\partial \storagecharge / \partial \demand \cdot \demand \Leftrightarrow 0$, implying that the when a storage charges power, it does not supply any demand. At these times, it the storage is a effective consumer retrieving power from producing assets. 
This leaves us with 
\begin{align}
     \allocatecapexstorage = \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \allocatestoragedispatch
\end{align}
where the dispatch allocation only assigns discharge power to demand $\demand$, thus
\begin{align}
    \allocatestoragedispatch \Leftrightarrow \pdv{\storagedispatch}{\demand} \demand
\end{align}
Note that this relation will break \cref{eq:cost_payback} as the payments to $r$ surpass the CAPEX by an amount $\remainingcost^E_r$. 
% \begin{align}
%     \capexstorage_r + \remainingcost^E_r = \sum_n \allocatecapexstorage
% \end{align} 
It is crucial to note that storage units perform a redistribution of money, and therefore distort the cost allocation. So, a certain share of what is allocated to the CAPEX of a storage is in another time step spent by the same storage in order to buy power from other assets. This effect scales with the amount of installed storage capacity.

We shortly annotate that it is possible to incorporate this redistribution into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}. Then, the derived payments that a storage unit $r$ has to pay to asset $i$ is given by $\cost_{r \rightarrow i}$. The sum of those payments due to $r$ will the sum up to $\remainingcost^E_r$.

% In a final step we can still allocate all cost to the demand $\demand$. However, this would require knowledge about when the power $\storagedispatch$ was initially charged: Given the amount of power $A_{r, t_{in} \rightarrow t_{out}}$ that is charged by storage $r$ at time step $t_{in}$ and discharged at $t_{out}$ ...

\subsection{Design Constraints}
\label{sec:design_constraints}
Power system modelling does rarely follow a pure Greenfield approach with unlimited capacity expansion. Rather, today's models are setting various constraints defining socio-political or  technical requirements. As mentioned before this will alter the equality of total cost and total revenue, \ie leads to $\remainingcost \ne 0$ in \cref{eq:total_revenue}. More precisely, each constraint $h_j$ (other then the nodal balance constraint) of the form 
\begin{align}
    h_j \left(\state, \capacity \right) - K < 0
    \label{eq:design_constraint}
\end{align}
where $K$ is any non-zero constant and $\capacity$ is collects all capacities of asset $i$, will result in a cost term contributing to $\remainingcost$ and alter \cref{eq:cost_decomposition} to 
\begin{align}
    \cost_i - \remainingcost_i = \sum_t \costfactor \, \state
\end{align}
The total allocatable cost associated to cost term $\cost$ is then given by 
\begin{align}
    \cost' = \sum_i  \cost_i - \remainingcost_i
    \label{altered_cost}
\end{align}
According to the nature of \cref{eq:design_constraint} and the corresponding $\remainingcost_i$, it is either larger, equal or lower than $\cost$.  
In the following we highlight two often used classes of designs constraints in the form of \cref{eq:design_constraint} and show how to incorporate them into the cost allocation.

\subsubsection{Capacity Expansion Limit}

In more realistic setups, generators, lines or other assets can only be built up to a certain limit. This might be due to land use restrictions or social acceptance problems. %In the following give a general solution how the cost allocation may handle this.  
However, when constraining the capacity $\capacity$  for a subset $I$ of assets to an upper limit $\capacityupper$, in the form of 
\begin{align}
    \capacity - \capacityupper \le 0 \resultsin{\muuppernom} \Forall{i \in I}
\label{eq:capacityexpansionmaximum},
\end{align}
the zero profit condition alters as soon as the constraint becomes binding. Then, the revenue of asset $i$ exceeds its total expenditures (OPEX + CAPEX). More precisely, the allocated CAPEX in \cref{tab:cost_allocation_map} will surpass the actual CAPEX of asset $i$ by the cost it has to pay for the scarcity, given by the absolute value of 
\begin{align}
    \scarcitycost_i = - \muuppernom \capacity \Forall{i \in I}
    \label{eq:scarcitycost}
\end{align}
% \begin{align}
%  \cost_i - \remainingcost_i =  \sum_t \costfactor \, \state \Forall{i}
%  \label{eq:capex_generation_duality_bf2}
% \end{align}

% In order to ensure that the allocated CAPEX sum up to the actual CAPEX at asset $i$, meaning that \cref{eq:cost_payback} holds, we adjust the CAPEX allocations to 
% \begin{align}
%     \allocatecost = \left(\dfrac{c_i}{c_i + \muuppernom}\right) \, \costfactor \, \allocatestate \Forall{i \in I}
% \end{align}
% with the cost factors $\costfactor$ for CAPEX given in \cref{tab:cost_allocation_map}. 
% The costs which consumers at $n$ have to pay for the scarcity impacting asset $i$ are given   
% \begin{align}
%     \allocatescarcitycost = \left(\dfrac{\muuppernom}{c_i + \muuppernom}\right) \, \costfactor \, \allocatestate \Forall{i \in I}
% \end{align}

The share of the CAPEX allocation, which results from the scarcity, is given by 
\begin{align}
    \allocatescarcitycost = \dfrac{\muuppernom}{c_i + \muuppernom} \, \allocatecost^I \Forall{i \in I}
\end{align}
where $\allocatecost^I$ denotes the CAPEX allocation presented above.

% have to be adjusted, as otherwise consumers would pay too much. As we show in ... updating the prices by a fix ratio 
% \begin{align}
%     \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppernom}\, \costfactor
%     \label{eq:adjustment_lowernom}
% \end{align}
% will . Note that $\muuppernom$ is a positive number, so the adjustment cannot increase the paid price.

\subsubsection{Brownfield Constraints}

In order to take already built infrastructure into account, the capacity $\capacity$ can be constrained to a minimum required capacity $\capacitylower$. Mathematically this translates to 
\begin{align}
    \capacitylower - \capacity  \le 0 \resultsin{\mulowernom} \hpad \forall{i \in I}
\label{eq:capacityexpansionminimum}
\end{align}
Again, such a setup alters the zero profit condition of asset $i$, as soon as the constraint becomes binding. 
In that case, asset $i$ does not collect enough revenue in order to match the CAPEX. The difference, given by 
\begin{align}
    \subsidycost_i = \mulowernom \capacity
    \Forall{i}
\end{align}
has to be subsidized by governments or communities. It is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost^I = \subsidycost_i $.

% Again, we can balance this effect out by updating the cost factor to 
% \begin{align}
%     \costfactor \rightarrow \costfactor +  \dfrac{\mulowernom \capacity}{\sum_t \state}
% \end{align}
% In contrast to \cref{eq:adjustment_lowernom} the price is not weighted by a factor but shifted for all time steps. This is required as 
% As $\mulowernom$ is a positive number, this will increase the price and therefore the revenue for asset $i$.

% In order to take this effect into account for the cost allocation, we update the cost factors in \cref{tab:cost_allocation_map} to

% \begin{align}
%     \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppergenerationnom}\, \costfactor
% \end{align}
% This ensures that the allocated CAPEX sum up to the actual CAPEX at asset $i$ (and \cref{eq:cost_payback} holds).

\section{Assumptions on Power Allocations}
% \section{\texorpdfstring{Extended Solution Space of $\boldsymbol{\slackk[m]}$}{Solution Space of the Slack}}
\label{sec:localizing_allocations}

The presented cost allocation suits for any type of topology and network setup. It may be extended by other classes of network assets or design constraints. But so far, the question of how the allocation $\allocatestate$ for generators $s$, lines $\ell$ and storages $r$ are defined was left open.

In a first step, we show that the needed information narrows down to peer-to-peer allocations between producing assets and consumers. %We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:state_allocation_constraint}). \\
Let $\injection[m]$ denote the net power injection at node $m$ and time $t$. It is given by 
\begin{align}
    \injection[m] = \sum_{i \in \{s, r\}} \incidenceasset[m] \, \state - \demand[m]
\end{align}
where $\incidenceasset[m]$ is 1 if asset $i$ is attached to bus $m$ and zero otherwise. Now, let $\ptdf$ denote a linear mapping between the injection $\injection[m]$ and the flow $\flow$, such that  
\begin{align}
 \flow  = \sum_m \ptdf[m]\, \injection[m]
 \label{eq:power_flow_equation}  
\end{align}
Usually, $\ptdf$ is given by the Power Transfer Distribution Factors (PTDF) which indicate the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $m$. For transport models or mixed AC-DC networks, these can calculated or extended using the formulation presented in \cite{hofmann_flow_2020-1}.

Using the two expression the derivative of the flow $\flow$ with respect to the demand can be expressed through 
\begin{align}
    \pdv{\flow}{\demand} \demand \Leftrightarrow  \allocateflow = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) 
 \label{eq:allocate_peer_to_allocate_flow},
\end{align}
where we collected all power produced at bus $m$ and consumed at bus $n$ in $\allocatepeer$, defined as
\begin{align}
    \allocatepeer = \sum_{i \in \{S,R\}} \incidenceasset[m] \, \allocatestate  
    % \nodalgeneration[m] = \sum_s \incidencegenerator[m] \, \generation + \sum_r \incidencestorage[m] \storagedispatch \Forall{n}
\end{align}
\Cref{eq:allocate_peer_to_allocate_flow} shows that $\allocateflow$ is fully determined through the peer-to-peer allocation $\allocatepeer$. In other words, we only need to know how much of the power produced at node $m$ is consumed at node $n$. With this information we may also derive the cost allocation for the transmission system. Unfortunately, the solution for $\allocatepeer$ is non-unique and requires further assumptions. Established flow allocation schemes approach this problem from different directions. Principally two options exist \textit{what} is allocated 
% 
\begin{enumerate}
\item gross power injections \label{gross}
\item net power injections \label{net}
\end{enumerate}
% 
Further it is important \textit{what assumptions} define the allocation, \ie what method is used to define the pairs of sources and sinks. The three suitable approaches we present here are
% 
\begin{enumerate}[label=\alph*., ref=\alph*]
\item Equivalent Bilateral Exchanges (EBE) \cite{galiana_transmission_2003} which assumes
that every producer supplies every consumer proportional to its share in the total consumption. \label{ebe} 
\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
\item Flow Based Market Coupling (FBMC) which uses zonal PTDF for allocating power within predefined regions. The interregional exchange is only allocating net power deficit or excess of the regions. \label{fbmc}
\end{enumerate}
% 
We show the mathematical formulation for all combinations \ref{ebe}\ref{gross} - \ref{fbmc}\ref{net} in \cref{sec:gross_ebe,sec:net_ebe,sec:gross_ap,sec:net_ap}.
% 
Principally, allocating net power instead of gross power leads to less P2P trades, as power from a bus $m$ with $\nodalgeneration[m] \le \nodaldemand[m]$ is not assigned to other buses, only to $m$.

When it comes to the methods, as literature has often pointed out, the EBE principal does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}. The AP allocation tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal. Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when following in the direction of $\flow$, ends up at bus $n$. In contrast, the FBMC allows to control the regions or market zones which are netted out in a first step. If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \nodalgeneration \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions. However, it relies on further assumptions such as the Generation Shift Keys determining the producing assets which are deployed for the inter-regional exchange. 
% 
\begin{figure*}[t]
    \centering
    \includegraphics[width=\linewidth]{example_network.png}
    \caption{Illustrative example of a 2-bus network with one optimized time step. Fixed prices and constraining values are given in the left box for each bus and the transmission line. Optimized values are given in the right boxes. Bus~1 has a cheaper operational price $o$, capital prices are the same for both. As both generator capacities are constraint to 100~MW, the optimization also deploys the generator at bus~2. The resulting electricity prices $\lambda$ are then a composition of all prices for operation and capital investments.}
    \label{fig:example_network}
\end{figure*}
%

From this point of view, we restrict the focus of this research to \textbf{net} injection allocation on the basis of \textbf{AP}, type \ref{net}\ref{ap}, as locality and no further dependency on external decisions are strong arguments for a transparent cost allocation. We emphasize that the AP based allocation does not imply traceability of the resulting flows, as it only determines the \textit{source}-\textit{sink} relations $\allocatepeer$. The flow allocation still follows the power flow equation given as shown in \cref{eq:allocate_peer_to_allocate_flow}.
In the following example, we illustrate the functionality of the cost allocation by an example of a small artificial network.
% % 
% 
\subsection{Numerical Example}
\label{sec:numerical_example}
% 
Consider a two bus system, depicted in \cref{fig:example_network}, with one transmission line and one generator per bus. Generator~1 (at bus~1) has a cheap operational price of 50 \euro/\megawatthour, generator~2 (at bus~2) has a expensive operational price of 200~\euro/\megawatthour. For both, capital investments amount 500~\euro/MW and the maximal capacity is limited to $\capacitygenerationupper$~=~100~MW. The transmission line has a capital price of 100~\euro/MW and no upper capacity limit. With a demand of 60~MW at bus~1 and 90~MW at bus~2, the optimization expands the cheaper generator at bus~1 to its full limit of 100~MW. The 40~MW excess power, not consumed at bus~1, flows to bus~2 where the generator is built with only 50~MW. \\
%  
\begin{figure*}[h!]
    \begin{subfigure}[c]{.495\linewidth}
    % NOTE: Net EBE and Net AP lead to the same result here. Due to time limits just take the ebe figures
    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1}
    \end{subfigure}
    \begin{subfigure}[c]{.495\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2}
    \end{subfigure}
    \caption{Power allocations of net power injections using the AP scheme, type \ref{net}\ref{ap}, in the example network, \cref{fig:example_network}. Consumers at for bus~1, figure (a), are, with 60~MW, totally supplied by the local generator and do not consume any imported power from bus 2. In contrast consumer at bus~2, figure (b), retrieve 40~MW from generator 1, induce a flow of 40~MW on the transmission line and consume 50~MW from the local generator.}
    \label{fig:example_allocation}
\end{figure*}
% 
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
    \caption{Full P2P cost allocation on the basis of power allocations shown in \cref{fig:example_allocation}. Consumers compensate OPEX and CAPEX of the generators they retrieve from, see \cref{fig:example_allocation}. As bus~1 is totally self-supplying, all it payment is assigned to the local generator.  As bus~2 imports power from bus~1 and thus induces a flow on line~1, it not only compensates local expenditures but also OPEX and CAPEX at bus~1 and CAPEX for the transmission.}
    \label{fig:example_payoff}
\end{figure}    
% 
% 
% 
\Cref{fig:example_allocation} shows the allocated transactions on the basis of allocation \ref{net}\ref{ap} for both buses 1~\&~2 separately. Note, the ``sum'' of the two figures give to the actual dispatch. The resulting P2P payments are given in \cref{fig:example_payoff}.

The left graph \cref{fig:example_allocation_bus1} shows that $d_1$ is with 60~MW totally supplied by the local production. Consequently consumers at bus~1 pay 3k~\euro~OPEX, which is the operational price of 50~\euro\, times the 60~MW. Further they pay 33k~\euro~ for the CAPEX at generator~1. Note that 3k~\euro~ of these account for the scarcity imposed buy the upper expansion limit $\capacitygenerationupper$. The rest makes up 60\% of the total CAPEX spent at generator~1, exactly the share of power allocated to $d_1$. Consumers at bus~1 don't pay any transmission CAPEX as no flow is associated with there demand. \\
The right graph \cref{fig:example_allocation_bus2} shows the power allocations to $d_2$. We see that 50~MW are self-supplied whereas the remaining 40~MW are imported from bus~1. Thus, consumers have the pay for the local OPEX and CAPEX as well as the proportional share of expenditures at bus~1 and the transmission system. As the capacity at generator~2 does not hit the expansion limit $\capacitygenerationupper$, no scarcity cost are assigned to it. The allocated CAPEX $\allocatecapexgeneration[2\rightarrow 2]$  compensates  the actual CAPEX of generator~2. Again. in the case of generator~1, 2\kk of the 22\kk CAPEX allocation are associated with the scarcity cost for generator~1. The payed congestion revenue of 4\kk is exactly the CAPEX of the transmission line, which directly reflects the zero-profit condition.   
% 
The sum of all values in the payoff matrix in \cref{fig:example_payoff} yield $\totalcost - \scarcitycost$, the total system cost minus the scarcity cost (which in turn is negative).
The sum of a column  yields the total revenue per the asset $i$. These values match their overall spending plus the cost of scarcity. The sum a row returns the total payment of consumers at bus $n$. For example the sum of payments of consumers at bus~1 is 36k~\euro. This is exactly the electricity price of 600~\euro/MW times the consumption of 60~MW, $\lambda_1 d_1$. \\
%

The fact that OPEX and CAPEX allocations are proportional to each other results from optimizing one time step only. This coherence breaks for larger optimization problems with multiple time steps. Then CAPEX allocation takes effect only for time steps in which one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.  \\
% The example shows that allocating net power injection leads to intuitive and transparent  cost allocations.

% In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments. The resulting payment given in \cref{fig:example_payoff_net_ebe} builds on the allocated power flow shown \cref{fig:example_allocation_net_ebe} in \cref{sec:example_plots}. As bus~2 does not produce excess power, none of its power production is assigned to bus~1 and thus no payment of bus~1 to bus~2 allocated. Neither has bus~1 to pay fee to the transmission system as it only exports power. So, consumers at bus~1 pay to its local generator. Bus~2 in contrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1. 
% % 
% Again the cumulative payments per bus meet the nodal spending $\lmp \, \demand$. The cumulative revenues per generator and transmission line meet the all CAPEX and OPEX. Note this gives the same result as when allocating net injection with the Average Participation \ref{net}\ref{ap}.

\section{Application Case}
% 
\begin{figure*}[t]
    \centering
    \includegraphics[width=\linewidth]{de50bf/network}
    \caption{Brownfield optimization of the German power system. The left side shows existent renewable capacities, matching the total capacity for the year 2017, which serve as lower capacity limits for the optimization. The right side shows the capacity expansion of renewables as well as installation of backup gas power plants. The effective CO$_2$ price is set to 120 \euro per tonne CO$_2$ emission.}
    \label{fig:network}
\end{figure*}
% 
We showcase the behavior of the cost allocation in a more complex system, by applying it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution. The model builds up on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.

We follow a brownfield approach where transmission lines can be expanded starting from today's capacity values, originally retrieved from the ENTSO-E Transmission System Map \cite{entso-e_entso-e_nodate}. Pre-installed wind and solar generation capacity totals of the year 2017 were distributed in proportion to the average power potential at each site excluding those with an average capacity factor of 10\%. Further, wind and solar capacity expansion are limited by land use restriction. These consider agriculture, urban, forested and protected areas based on the CORINE and NATURA2000 database \cite{corine2012,natura2000}. Pumped Hydro Storages (PHS) and Run-of-River power plants are fixed to today's capacities with no more expansion allowed. Additionally, unlimited expansion of batteries and H$_{2}$-storages and Open-Cycle Gas Turbines (OCGT) are allowed at each node. 
% 
% 
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{de50bf/average_price}
    \caption{Average electricity price $\averagelmp$ per region as a result from the optimization of the German power system. Regions in the middle and south of Germany have high prices whereas electricity in the North with a strong wind, transmission and OCGT infrastructure is cheaper.}
    \label{fig:average_price}
\end{figure}
We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$ which, for OCGT, adds an effective price of 55 \euro/\megawatthour (assuming a gross emission of 180~kg/MWh and an efficiency of 39\%). All cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are summarized in detail in \cref{tab:cost_assumptions}.

The optimized network is shown in  \cref{fig:network}. On the left we find the lower capacity bounds for renewable generators and transmission infrastructure, on the right the capacity expansion for generation, storage and transmission are displayed. The optimization expands solar capacities in the south, onshore and offshore wind in the upper north and most west. Open-Cycle Gas Turbines (OCGT) are build within the broad middle of the network. Transmission lines are amplified in along the North-South axis, including one large DC link, associated with the German S\"ud-Link, leading from the coastal region to the South-West. 
The total annualized cost of the power system roughly sums up to 42 billion \euro.
%

\Cref{fig:average_price} displays the average electricity price $\lmp$ per region. We observe a relatively strong gradient from south (at roughly 92 \euro/MWh) to north (80 \euro/MWh). Regions with with little pre-installed capacity and capacity expansion, especially for renewables, tend to have higher prices. The node with the lowest LMP in the upper North-West, stands out through high pre-installed offshore capacities. \\

The costs are allocated according to type \ref{net}\ref{ap}. For allocating flows on  DC-lines we use the method described in \cite{hofmann_flow_2020-1}.
In \cref{fig:total_cost} we show the total of all ``allocatable'' costs $\cost'$. Due to the design constraints, \cref{sec:design_constraints}, and storage units, \cref{sec:storages}, these are different from the original cost terms $\cost$. The figure also includes costs for scarcity $\scarcitycost$ and subsidy  $\subsidycost$. Note that the sum of all contributions in \cref{fig:total_cost} equals the total cost $\totalcost$.

\begin{figure}
    \centering
    \includegraphics[width=0.75\linewidth]{de50bf/total_costs}
    \caption{Total allocated payments of the system. }
    \label{fig:total_cost}
\end{figure}

The largest proportion of the payments is associated with CAPEX for generators, transmission system and storage units in decreasing order. 
Taking different technologies into account, we observe a fundamental difference between the controllable OCGT  and the variable renewable resources: As shown in detail in \cref{fig:capex_duration_curve}, more than half of all investments in OCGT is determined in one specific hour. At this time (morning, end of February), the system hits the highest mismatch between low renewable power potential and high demand. The necessity for backup generators manifests in high allocation of CAPEX and consequently high LMP. With a few exceptions in the South West and East, all consumers receive power from OCGT at this time, thus all pay high amounts for the needed backup infrastructure (operational state of the system is in detailed shown in \cref{fig:operation_high_expenditure}). Note that this is the most extreme event, which ensures backup infrastructure for other inferior extreme events. The total CAPEX allocation for OCGT infrastructure is depicted in \cref{fig:ocgt_capex}. This  correlates with our findings of the extreme event. \\
Contrary to this, CAPEX for renewable infrastructure are allocated evenly throughout several thousands of hours. As for onshore and offshore wind farms, the produced power deeply penetrates the network, see \cref{fig:power_mix}, thus it is not only local, but also remote consumers which cover the CAPEX, see \cref{fig:capex_price}. This in turn benefits local consumers, which profit from the cheap power which explains why these regions end up with a low average LMP.\\

Together with the emission cost $\emissioncost$, the total OPEX $\opex$ amount around 16 billion \euro. As to expect, 99.97\% are dedicated to OCGT alone, as these have by far the highest operational price. For a detailed regional distribution of payments per MWh and the resulting revenues for generator, see \cref{fig:opex_price,fig:emission_cost}. For OCGT, the spatial allocation of the OPEX clearly differs from  the CAPEX allocation. This results from the fact that during ordinary demand peaks, it is rather local OCGT generators which serve as backup generators. In the average power mix per region, see \cref{fig:power_mix}, we observe that regions with strong OCGT capacities predominately have high shares of OCGT power. The OPEX allocations for renewables play with 0.2 \euro/MWh in average an inferior role. \\

\begin{figure*}
    \centering
    \begin{subfigure}[c]{.6\linewidth}
    \includegraphics[width=\linewidth]{de50bf/ptpf_net_to_lowest-lmp}
    \end{subfigure}
    \hspace{-.2151\linewidth}
    \begin{subfigure}[c]{.6\linewidth}
    \includegraphics[width=\linewidth]{de50bf/ptpf_net_to_highest-lmp}
    \end{subfigure}
    \caption{Comparison of payments of the node with the \textbf{lowest LMP (left)} and the node with the \textbf{highest LMP (right)}. The region of the paying bus is colored in dark blue. The circles indicate where to which bus and technology combined OPEX and CAPEX payments. Further the thickness of the lines indicates the dedicated amount of payments. The cheap prices in the North...  }
    \label{fig:direct-allocation}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{de50bf/locality_all_ptpf_net}
    \caption{Average distance between payer and receiver for different technologies and shares of the total production.}
    \label{fig:locality}
\end{figure}

The negative segment in \cref{fig:total_cost} is associated with scarcity costs $\scarcitycost$, caused by land use constraints for renewables and the transmission expansion limit. These sum up approximately to 7.5 b\euro. Note again, that according to \cref{eq:scarcitycost} this term is negative and is part of the allocated CAPEX payments. It translates to the cost that consumers pay ``too much'' for assets limited in their capacity expansion. In the real world this money would be spent for augmented land costs or civic participation in the dedicated areas. In \cref{fig:scarcity_price} we give a detailed insight of how the scarcity cost manifest in the average cost per consumed MWh. The scarcity for wind and solar is relatively low and impact the average price at roughly 2\euro/MWh. It primarily affects, different to the pure CAPEX allocation, regions in close vicinity. Remarkably, the scarcity cost per MWh for run-of-river power plants amounts up to 16\euro. This high impact is due to the steady power potential from run-off water and the strong limitation of capacity expansion. However, as these power plants in particular are already amortized, the scarcity cost should be reconsidered and removed from a final cost allocation.

A high influence on the price comes from the scarcity of transmission expansion. Right beside regions with high wind infrastructure, it occasions costs higher than MWh 4\euro/MWh (maximal 8\euro/MWh), see \cref{fig:branch_scarcity_price}. As to expect, the constraint mainly suppresses transmission expansion along the north-south axis.

The last cost term in \cref{fig:total_cost}, is  caused by lower capacity constraints for pre-existing assets. These violate the optimal design and are not compensated through the nodal payments $\demand \, \lmp$. Most of these ``non-allocatable'' costs account for Pumped Hydro Storage, onshore wind, offshore wind and the transmission system, see \cref{fig:subsidy} for further details. Again, as most of the named assets are already amortized (PHS, transmission system), the subsidy should be reconsidered in a final cost allocation.

In \cref{fig:direct-allocation} we compare the total cost allocation of the region with the lowest average LMP (left side) against the one of the region with the highest LMP (right side). The region with the cheapest electricity is in the Northwest. We see that this bus is fairly independent of investments in the transmission system, as it is supplied by steady, cheap offshore power. Moreover, the CAPEX of the local offshore wind farms are partly payed by subsidies, see \cref{fig:subsidy}. Only a small share of the payments is allocated to remote OCGTs, resulting from the extreme event mentioned earlier.

In contrast the regions with the highest LMP, spends most its money to local OCGT operations, emission cost and the transmission system. Its payments to  onshore and offshore wind infrastructure are low despite a third of its supply comes from wind power. Instead, a lot of its payments are assigned to the transmission infrastructure. Hence, the wind power supply at this region is not restricted by exhausted wind power resources but by bottlenecks in the transmission system.

\section{Conclusion}

In this work we presented a new cost-allocation scheme based on peer-to-peer dispatch allocations from asset to consumer. Starting from an optimized long-term investment model, we were able to decompose single cost terms into contributions of single consumers. For three typical classes of assets, namely generators, transmission lines and storage units, we showed how operational prices and shadow prices must be weighted with the dispatch allocation in order to allocate all system costs. Further we highlighted the impact of additional design constraints, that is constraints with a non-zero constant on the right hand side, such as lower and upper capacity expansion limits. These alter the locational marginal price and therefore distort the cost allocations. In the case of upper capacity expansion limits, this leads to an additional charge for consumers which have to compensate for \eg land use restrictions. For lower capacity investment bounds, the cost have to be subsidized by the overall system operator or the government, as those cannot be allocated.
\dots
\clearpage
\appendix

\section{Network Optimization}

\renewcommand\theequation{\thesection.\arabic{equation}}
\setcounter{equation}{0}

\renewcommand\thefigure{\thesection.\arabic{figure}}    
\setcounter{figure}{0}

\subsection{LMP from Optimization}
The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL). Alternatively, we can the demand $\demand$ has to be supplied by the attached assets,  
\begin{align}
    \sum_i \incidenceasset \state  &=  \demand 
    % \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0
     \resultsin{\lmp} \Forall{n,t}
    \label[constraint]{eq:nodal_balance_lin}
\end{align}
where $\incidenceasset$ is +1 if $i$ is attached to $n$ and a positive operation $\state$ delivers power to $n$, -1 if is attached to $n$ and a positive operation retrieves power from $n$ and zero else (note that for lines this results in the negative of the conventional Incidence Matrix).  
The shadow price of the nodal balance constraint mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay.\\

\subsection{Full Lagrangian}
\label{sec:full_lagrangian}
The Lagrangian for the investment model can be condensed to the following expression

\begin{align}
\notag
 \lagrangian\left(\state, \capacity, \lmp, \mu_j \right) =& 
 \;\;\; \sum_{i,t} o_i \state + \sum_{i} c_i \capacity  \\
\notag
&+  \sum_{n,t} \lmp \left( \demand - \sum_i \incidenceasset \, \state \right) \\
&+ \sum_j \mu_j \, h_j \left(\state, \capacity \right)
\label{eq:full_lagrangian}
\end{align}
where $h_j \left(\state, \capacity \right)$ denotes all inequality constraints attached to $\state$ and $\capacity$.
% \begin{align}
% \notag
% & \lagrangian\left(\generation, \flow, \capacitygeneration, \capacityflow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
% \notag
% & \;\;\; \sum_{n,s} \capitalpricegeneration \capacitygeneration + \sum_{n, s, t} \operationalpricegeneration \generation + \sum_{\ell} \capitalpriceflow \, \capacityflow  \\
% \notag
% &+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \incidencegenerator\, \generation +  \demand  \right)  \\ 
% \notag
% &+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \reactance \, \flow  \\
% \notag
% &+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacitygeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
% &+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityflow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityflow \right)     
% \label{eq:full_lagrangian}
% \end{align}
% 
% where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables. 
% 
In order to impose the Kirchhoff Voltage Law (KVL) for the linearized AC flow, the term 
\begin{align}
    \sum_{\ell,c,t} \cycleprice \, \cycle \, \reactance \, \flow 
\end{align}
can be added to $\lagrangian$, with $\reactance$ denoting the line's impedance and $\cycle$ being 1 if $\ell$ is part of the cycle $c$ and zero otherwise.

The global maximum of the Lagrangian requires stationarity with respect to all variables:
\begin{align}
    \pdv{\lagrangian}{\state} = \pdv{\lagrangian}{\capacity} = 0    
\end{align}

\subsection{Zero Profit Generation}
\label{sec:zero_profit_generation}
\Cref{eq:upper_generation_capacity_constraint,eq:lower_generation_capacity_constraint}, which yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$, imply the complementary slackness,
\begin{align}
\muuppergeneration \left( \generation - \generationpotential \, \capacitygeneration \right)  &= 0  \Forall{n,s,t} 
\label{eq:complementary_slackness_upper_generation} \\
\mulowergeneration  \, \generation &= 0 \Forall{n,s,t}
\label{eq:complementary_slackness_lower_generation} 
\end{align}

The stationarity of the generation capacity variable leads to 
\begin{align}
\pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, 
\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
\label{eq:capex_generation_duality}
\end{align}
and the stationarity of the generation to 
\begin{align}
\pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
\operationalpricegeneration =  \incidencegenerator \, \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
\end{align}

Multiplying both sides of \cref{eq:capex_generation_duality} with $\capacitygeneration$ and using \cref{eq:complementary_slackness_upper_generation} leads to 
\begin{align}
 \capitalpricegeneration \, \capacitygeneration  = \sum_t \muuppergeneration \, \generation 
 \label{eq:capital_price_generation_sum}
\end{align}
The zero-profit rule for generators is obtained by multiplying \cref{eq:opex_duality} with $\generation$ and using \cref{eq:complementary_slackness_lower_generation,eq:capital_price_generation_sum} which results in 
\begin{align}
  \capitalpricegeneration \, \capacitygeneration + \sum_t \operationalpricegeneration \generation = \sum_t \lmp \incidencegenerator \generation
\end{align}
It states that over the whole time span, all OPEX and CAPEX for generator $s$ (left hand side) are payed back by its revenue (right hand side).

\subsection{Zero Profit Transmission System}
\label{sec:zero_profit_flow}

The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the transmission capacity in positive or negative direction, i.e. \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding. For flows below the thermal limit, the complementary slackness 
\begin{align}
\muupperflow \left( \flow - \capacityflow \right)  &= 0 \Forall{\ell,t}
\label{eq:complementary_slackness_upper_flow} \\
\mulowerflow \left( \flow - \capacityflow \right) &=  0 \Forall{\ell,t}
\label{eq:complementary_slackness_lower_flow} 
\end{align}
sets the respective KKT to zero.

The stationarity of the transmission capacity to
\begin{align}
\pdv{\lagrangian}{\capacityflow} = 0 \,\, \rightarrow \,\, 
\capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
\label{eq:capexFlow_duality}
\end{align}
and the stationarity with respect to the flow to
\begin{align}
    0 &= \pdv{\lagrangian}{\flow}  \\ 
    0 &= - \sum_n \incidence \lmp  + \cycleprice \cycle \reactance  - \muupperflow + \mulowerflow \Forall{n,s} \label{eq:opex_flow_duality}
\end{align}

When multiplying \cref{eq:capexFlow_duality} with $\capacityflow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain 
\begin{align}
 \capitalpriceflow \, \capacityflow = \sum_t \left( \muupperflow - \mulowerflow \right)  \, \flow
 \label{eq:capital_price_flow_sum}
\end{align}
Again we can use this to formulate the zero-profit rule for transmission lines. We multiply \cref{eq:opex_flow_duality} with $\flow$, which finally leads us to 
\begin{align}
\capitalpriceflow \, \capacityflow = - \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \reactance\, \flow 
\end{align}
% Watch out Incidence = - ordinary Incidence !!
It states that the congestion revenue of a line (first term right hand side) reduced by the cost for cycle constraint exactly matches its CAPEX.

\subsection{Zero Profit Storage Units}
\label{sec:zero_profit_storage_units}

For an simplified storage model, the upper capacity $\capacitystorage$ limits the discharging dispatch $\storagedispatch$, the storing power $\storagecharge$ and state of charge $\storagesoc$ of a storage unit $r$ by 
\begin{align}
    \storagedispatch - \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragedispatch} \\
    \storagecharge - \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragecharge} \\
    \storagesoc - h_r \, \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragesoc}
\end{align}
where we assume a fixed ratio between dispatch and storage capacity of $h_r$. 
% From stationarity we obtain 
% \begin{align}
%     \notag
%     \pdv{\lagrangian}{\capacitystorage} &= 0 \\
%     \capitalpricestorage &= \sum_t \left( \muupperstoragedispatch + \muupperstoragecharge + h_r \muupperstoragesoc  \right)
% \end{align}
The state of charge must be consistent throughout every time step according to what is dispatched and stored, 
\begin{align}
    \notag
    \storagesoc - \efficiencysoc \storageprevioussoc - \efficiencycharge \storagecharge &+ (\efficiencydispatch)^{-1} \storagedispatch = 0 \\
    &\resultsin{\mustateofcharge} \Forall{r,t}
\end{align}

We use the result of Appendix B.3 in \cite{brown_decreasing_2020} which shows that a storage recovers its capital (and operational) costs from aligning dispatch and charging to the LMP, thus 
\begin{align}
    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage \left(\storagedispatch - \storagecharge \right) 
\end{align}
The stationarity of the dispatched power leads us to  
\begin{align}
    \notag
    \pdv{\lagrangian}{\storagedispatch} &= 0 \\
    \operationalpricestorage -  \lmp \, \incidencestorage - \mulowerstoragedispatch + \muupperstoragedispatch + (\efficiencydispatch )^{-1} \mustateofcharge &= 0  
    \label{eq:stationarity_storagedispatch}
\end{align}
which we  can use to define the revenue which compensates the CAPEX at $r$, 
\begin{align}
    \notag
    \capitalpricestorage \, \capacitystorage = \sum_t \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \storagedispatch \\
    - \sum_t \lmp \incidencestorage  \storagecharge \Forall{r} 
\end{align}

% With this constraint we can derive the following stationarities:
% \begin{align}
%     \notag
%     \pdv{\lagrangian}{\storagedispatch} &= 0 \\
%     \operationalpricestorage + \lmp + \mulowerstoragedispatch - \muupperstoragedispatch - (\efficiencydispatch )^{-1} \mustateofcharge &= 0  
%     \label{eq:stationarity_storagedispatch}
% \end{align}
% \begin{align}
%     \notag
%     \pdv{\lagrangian}{\storagecharge} &= 0 \\
%     - \lmp + \mulowerstoragecharge - \muupperstoragecharge + \efficiencycharge \mustateofcharge &= 0 
%     \label{eq:stationarity_storagecharge}
% \end{align}
% \begin{align}
%     \notag
%     \pdv{\lagrangian}{\storagesoc} &= 0 \\
%     \mulowerstoragesoc - \muupperstoragesoc - \mustateofcharge + \efficiencysoc \munextstateofcharge &= 0 
%     \label{eq:stationarity_storagesoc}
% \end{align}

\subsection{Emission Constraint}

Imposing an additional CO$_2$ constraint limiting the total emission to K,  
\begin{align}
\sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionprice} 
\label[constraint]{eq:co2_constraint}
\end{align}
with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionprice$ in \euro/tonne-CO$_2$. 
% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
% 
% \begin{align}
% \capitalpricegeneration \, \capacitygeneration + \sum_{t} \operationalpricegeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionprice \right)  \, \generation \Forall{n,s} 
% \label{eq:non_profit_generator_emission}
% \end{align}
% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
\begin{align}
\operationalpricegeneration \rightarrow \operationalpricegeneration + \emission \, \emissionprice \label[relation]{eq:shift_opex_by_emission_cost}
\end{align}

% \subsection{Brownfield Optimization and Capacity Restrictions}

% Constraining the capacities $\capacitygeneration$  for a subset $S$ of generators to lower or upper limits in the form of
% \begin{align}
%     \capacitygenerationlower - \capacitygeneration \le 0 \resultsin{\mulowergenerationnom} \hpad \forall{s \in S} \label{eq:capacityexpansionminimum}\\
%     \capacitygeneration - \capacitygenerationupper \le 0 \resultsin{\muuppergenerationnom} \hpad \forall{s \in S}
% \label{eq:capacityexpansionmaximum}
% \end{align}
% alters the objective value as soon as one of those become bounding. 
% The complementary slackness for both are 
% \begin{align}
%     \muuppergenerationnom \left( \capacitygeneration - \capacitygenerationupper \right) =0 \Forall{s \in S} \\
%     \mulowergenerationnom \left( \capacitygeneration - \capacitygenerationlower \right) =0 \Forall{s \in S}
% \end{align}

% The CAPEX paybacks for generators and transmission lines
% \cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
% \begin{align}
% \pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, \\
% \capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{s \in S}
% \label{eq:capex_generation_duality_bf}
% \end{align}
% for generators. Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacitygeneration$ leads us to 
% \begin{align}
%  \capexgeneration_s - \remainingcost_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
%  \label{eq:capex_generation_duality_bf2}
% \end{align}
% where we define the cost resulting from the capacity expansion limits as
% \begin{align}
%     \remainingcost_s = \left(\mulowergenerationnom - \muuppergenerationnom \right) \capacitygeneration
%     \Forall{s}
% \end{align}
% As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \ge 0$, the latter can be a net positive or net negative cost term.

% Multiplying \cref{eq:opex_duality}, which is still valid, with $\generation$ and inserting \cref{eq:capex_generation_duality_bf2} will bring us to the zero profit rule for generators with capacity expansion limits, 
% \begin{align}
%     \capexgeneration_s + \opexgeneration_s  - \remainingcost_s = \sum_t \lmp \incidencegenerator \generation \Forall{s \in S}
% \end{align}
% We see that the revenue of generator $s$ will not exactly pay back its full CAPEX and OPEX. The exogenous constraint shifts the zero-profit equations such that some of the expenditures for $s \in S$ cannot directly be allocated to consumers. In order to take this effect into account for the cost allocation, we update the shadow price $\muuppergeneration$ in \cref{tab:cost_allocation_map}, to

% \begin{align}
%     \muuppergeneration \rightarrow \dfrac{\capitalpricegeneration}{\capitalpricegeneration - \mulowergenerationnom + \muuppergenerationnom}\, \muuppergeneration
% \end{align}
% By doing so we ensure that the allocated CAPEX sum up to the actual CAPEX at generator $s$, thus 
% \begin{align}
%     \capexgeneration_s = \sum_{n,t} \allocatecapexgeneration \Forall{s \in S}
% \end{align}

% Doing likewise with a subset $L$ of transmission lines,
% \begin{align}
% \capacityflow \ge \capacityflowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
% \capacityflow \le \capacityflowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
% \end{align}
% results in a shifted zero-profit rule for transmission line in the form 
% \begin{align}
%     \capitalpriceflow \, \capacityflow - \remainingcost_\ell = \sum_{n,t}\allocatecapexflow \Forall{\ell \in L}
% \end{align}
% where $\remainingcost_\ell$ is given by 
% \begin{align}
%     \remainingcost_\ell = \left(\mulowerflownom - \muupperflownom \right) \, \capacityflow
%     \Forall{\ell \in L}
% \end{align}

% % \begin{align}
% %     \muupperflownom \left( \capacityflow - \capacityflowUpper \right) =0 \\
% %     \mulowerflownom \left( \capacityflow - \capacityflowLower \right) =0 \\
% % \end{align}

% % \begin{align}
% %     \pdv{\lagrangian}{\capacityflow}  = 0 \,\, \rightarrow \,\, \\
% %     \capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
% %     \label{eq:capex_flow_duality_bf}
% %     \end{align}
% %     for transmission lines.
% % Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacitygeneration - \capacitygenerationlower \right) $
% % \begin{align}
% %   \capitalpricegeneration \, \capacitygeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacitygenerationlower \right) + \muuppergenerationnom \left( \capacitygenerationupper - \capacitygenerationlower \right) \Forall{n,s}
% % \end{align}

\section{Allocation Schemes}
\subsection{Allocating Gross Injections with EBE}
\label{sec:gross_ebe}

The allocation of gross generation to demands $\demand$ is straightforwardly obtained by a proportional distribution of the generation, \ie

\begin{align}
    \allocategeneration = \dfrac{\generation}{\sum_s \generation} \demand 
\end{align}

\subsection{Allocating Net Injections with EBE}
\label{sec:net_ebe}

Allocating net power injections using the EBE methods leads to the same result as the Marginal Participation (MP) \cite{rudnick_marginal_1995}  algorithm when allocating to consumers only, see \cite{hofmann_flow_2020-1} for further insight. We calculate it by setting 
\begin{align}
\allocatepeer &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
\label{eq:mp_slack}
\end{align}
where 
\begin{itemize}
%  \item $\injection = \left( \nodalgeneration - \nodaldemand \right) $ denotes the nodal injection
\item $\netproduction = \text{min}\left( \nodalgeneration - \nodaldemand , 0 \right) $ denotes the nodal net production 
\item $\netconsumption = \text{min}\left( \nodaldemand  - \nodalgeneration, 0 \right)$ denotes the nodal net consumption
\item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the denotes  nodal self-consumption. That is the power generated and at the same time consumed at node $n$ and 
\item $\gamma_t = \left( \sum_n \netproduction\right) ^{-1} = \left( \sum_n \netconsumption\right) ^{-1}$ is the inverse of the total injected/extracted power at time $t$.
\end{itemize}

The allocation $\allocategeneration$ from generator $s$ to $n$, is given by multiplying $\allocatepeer$ with the nodal share $\generation / \nodalgeneration$.

\subsection{Allocating Net Power using AP}
\label{sec:net_ap}

\newcommand{\incidenceM}{K}
\newcommand{\flowM}{f}
\newcommand{\injectionM}{p}
\newcommand{\slackM}{k}
\newcommand{\DirectedIncidence}{\mathcal{K}}
\newcommand{\InverseAPInjection}{\mathcal{J}}
\newcommand\diag[1]{\operatorname{diag}\left(#1\right)}

Allocating net injections using the AP method is derived from \cite{achayuthakan_electricity_2010}. In a lossless network the downstream and upstream formulations result in the same P2P allocation which is why we restrict ourselves to the downstream formulation only. In a first step we define a time-dependent auxiliary matrix $\InverseAPInjection_t$ which is the inverse of the $N\times N$ with directed power flow $m \rightarrow n$ at entry $(m, n)$ for $m \ne n$ and the total flow passing node $m$ at entry $\left( m, m\right)$ at time step $t$. Mathematically this translates to

\begin{align}
\InverseAPInjection_t = \left( \diag{\injectionM^+} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
\end{align}
where $\DirectedIncidence^-$ is the negative part of the directed Incidence matrix $\DirectedIncidence_{n,\ell} = \text{sign}\left( \flow \right)  \incidence$. Then the distributed slack for time step $t$ is given by
\begin{align}
\allocatepeer = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
\end{align}

\subsection{Allocating Gross Power using AP}
\label{sec:gross_ap}

We use the same allocation as in \cref{sec:net_ap} but replace the net nodal production $\netproduction$ by the gross nodal production $\nodalgeneration$ which leads to  
\begin{align}
\InverseAPInjection_t = \left( \diag{g} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
\end{align}
The distributed slack is for time step $t$ is then given by
\begin{align}
\allocategeneration[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
\end{align}

% 
\section{Working Example}
The following figures contain more detailed information about the peer-to-peer cost allocation. The cost or prices payed by consumers are indicated by the region color. The dedicated revenue is displayed in proportion to the size of cycles (for assets attached to buses) or to the thickness of transmission branches.    
\begin{figure*}
    \centering
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/one_port_investment_cost_total}
        \subcaption{All production and storage technologies}
        \label{fig:total_capex}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/onwind_one_port_investment_cost}
        \subcaption{Onshore Wind}
        \label{fig:onshore_capex}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/solar_one_port_investment_cost}
        \subcaption{Solar}
        \label{fig:solar_capex}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/OCGT_one_port_investment_cost}
        \subcaption{OCGT}
        \label{fig:ocgt_capex}
    \end{subfigure}
    \caption{Average \textbf{CAPEX allocation} per MWh, $\sum_t \allocatecapexgeneration / \sum_t \demand$ and $\sum_t \allocatecapexstorage / \sum_t \demand$, for all production and storage assets (a), onshore wind (b), solar (c) and OCGT (d). Average Allocated CAPEX per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
    \label{fig:capex_price}
\end{figure*}

\begin{table*}[h]
    \centering
    \input{../tables/de50bf_ptpf_net/prices}    
    \caption{Operational and capital price assumptions for all type of assets used in the working example. The capital price for transmission lines are given in [k\,\euro/MW/km]. The cost assumptions are retrieved from the PyPSA-EUR model \cite{horsch_jonas_pypsa-eur_2020}.}
    \label{tab:cost_assumptions}
\end{table*}

\begin{figure}
    \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/one_port_operational_cost_total}
    \caption{Average \textbf{OPEX allocation} per consumed MWh, $\sum_t \allocateopex/\sum_t \demand$. The effective prices for OPEX are indicated by the color of the region, the size of the circles are set proportional to the revenue per regional generators and storages. As OCGT is the only allowed fossil based technology, the drawn allocation is proportional to OPEX allocation of OCGT generators.}
    \label{fig:opex_price}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/co2_cost_total}
    \caption{Average \textbf{allocated emission cost}, $\sum_t \allocateemissioncost$, per consumed MWh. The effective prices are indicated by the color of the region, the size of the circles are set proportional to the revenue per regional generators.}
    \label{fig:emission_cost}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{de50bf/power_mix_ptpf_net}
    \caption{Average power mix per region calculated by Average Participation \ref{ap}. Coastal regions are mainly supplied by local offshore and onshore wind farms. Their strong power injections additionally penetrate the network up to the southern border. In the middle and South, the supply is dominated by a combination of OCGT and solar power.}
    \label{fig:power_mix}
\end{figure}

\begin{figure}
    \vspace{2cm}
    \includegraphics[width=\linewidth]{de50bf/subsidy}
    \caption{Total costs for subsidy $\subsidycost$ resulting from lower capacity expansion bounds (brownfield constraints). The figure shows the built infrastructure that does not gain back its CAPEX from its market revenue, but is only build due to lower capacity limits.}
    \label{fig:subsidy}
\end{figure}

\begin{figure*}
    \includegraphics[width=\linewidth]{de50bf/capex_duration_curve}
    \caption{Duration curve of the CAPEX allocation for production and storage technologies. Hours are sorted by the total amount of allocated expenditure. With 2.7 bn~\euro\, the first value pushes investments extraordinarily high. Due to low renewable potentials, it is dominated by CAPEX for OCGT which receives 92\% of the payments. This hour alone occasions about the half of all OCGT CAPEX. \Cref{fig:capex_duration_curve} gives a detailed picture of the operational state at this time-step. The following 7000 time-steps are dominated by revenues for onshore wind and reveal a rather even distribution. In hours of low CAPEX allocation (after the second drop) spendings for OCGT start to increase again. These time-steps however play a minor role.}
    \label{fig:capex_duration_curve}
\end{figure*}

\begin{figure}
    \includegraphics[width=\linewidth]{de50bf/operation_high_expenditure}
    \caption{Production, flow and consumption in the system at the hour with the highest allocated expenditures. The size of the circles are proportional to the power production at a node. Size of arrows are proportional to the flow on the transmission line. The depicted hour corresponds to the first value in the duration curve in \cref{fig:capex_duration_curve}.}
    \label{fig:operation_high_expenditure}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/branch_scarcity_cost_total}
    \caption{Average \textbf{allocated transmission scarcity cost} per consumed MWh, $\sum_t \scarcitycost_{n \rightarrow \ell, t} / \sum_t \demand $. This scarcity cost results from the upper transmission expansion limit of 25\%. The costs are indicated by the regional color.  The lines are drawn in proportion to revenue dedicated to scarcity cost. }
    \label{fig:branch_scarcity_price}
\end{figure}

\begin{figure*}
    \centering
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/offwind-dc_generator_scarcity_cost}
        \subcaption{Offshore Wind}
        \label{fig:offwind-dc_generator_scarcity_cost}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/onwind_generator_scarcity_cost}
        \subcaption{Onshore Wind}
        \label{fig:onwind_generator_scarcity_cost}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/solar_generator_scarcity_cost}
        \subcaption{Solar}
        \label{fig:solar_generator_scarcity_cost}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/ror_generator_scarcity_cost}
        \subcaption{Run-of-River}
        \label{fig:ror_generator_scarcity_cost}
    \end{subfigure}
    \caption{Average \textbf{allocated scarcity cost} per consumed MWh, $\sum_t \allocatescarcitycost / \sum_t \demand$. These cost result from land use restricitions for offshore wind, onshore wind,  solar, run-of-river. The cost per MWh are indicated by the color of a region. The revenue per production asset is given by the size of the circle at the corresponding bus.}
    \label{fig:scarcity_price}
\end{figure*}

\clearpage
\printbibliography

\end{document}